---
title: "Maternal Health and the UPR"
params:
  country_name: "Bahrain"
  upr_all: NULL
  rec_plot: NULL
  mmr_map_plot: NULL
  bbox_SUR_region_dynamic: NULL
  bbox_selected_SUR: NULL
  sur_area: NULL
documentclass: article
# classoption: twocolumn
output: 
  pdf_document:
    includes:
      in_header: preamble.tex
    keep_tex: true
    latex_engine: xelatex
geometry: 
  - top=2cm
  - bottom=0.5cm
  - left=0.2cm
  - right=0.2cm
papersize: a4
date: "2025-09-30"
---
```{r setup, include=FALSE}
# This chunk loads the necessary packages for the report
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.align='center', out.width="100%")
pacman::p_load(
  here,
  dplyr, forcats, ggplot2, magrittr, readr, readxl, stringr, tibble, tidyr, lubridate,
  rlang,
  ggtext, 
  patchwork,
  kableExtra,
  # apyramid,
  sf
)

# sdg_data <- readRDS(here("data", "SDG_data_enhanced.rds")) |> droplevels() |> 
#   mutate(response_upr = fct_recode(response_upr, "Noted" = "Noted/Other"))
sdg_data <- readRDS(here("output", "UHRI_UPR_enhanced.rds")) |> 
  mutate(
    response_upr_2 = response_upr,
    response_upr = fct_recode(response_upr, 
                              "Noted" = "Partially supported"),
    response_upr = fct_relevel(response_upr, "Noted")) |> 
  droplevels()
state_geo <- readRDS(here("output", "state_geo_enhanced.rds"))
nearest_neighbors_list <- readRDS(here("output", "nearest_neighbors_list.rds"))
source(here("code", "theme_labels.R"))
iso3_SUR <- state_geo |> filter(country==params$country_name) |> pull(iso3)

# Loop through API-generated files
for (file_name in list.files(path = here("data", "API_data"), pattern = "\\.rds$")) {
  object_name <- gsub("\\.rds$", "", file_name)
  assign(object_name, readRDS(here("data", "API_data", file_name)))
}

# UPR plot data
health_rec_C <- sdg_data |> 
  filter(state_under_review == params$country_name) |> 
  droplevels() |>
  group_by(cycle, state_under_review) |>
  count(health_related, .drop = FALSE) |> 
  group_by(cycle) |>
  mutate(
    n_tot = sum(n),
    perc = (n / n_tot) * 100,
    perc = case_when(
      health_related == "Other" ~ "",
      .default = paste0(sprintf("%1.0f", perc), "%")
    )
  ) |> ungroup()

# Life expectancy
life_expectancy <- e01dt |> 
  left_join(state_geo |> sf::st_drop_geometry() |> select(iso3,iso2, loc_id),
            join_by(country_code==loc_id)) |> 
  filter(
    !is.na(iso3),
    year>=2005
  );rm(e01dt)

pop_wpp <- pop1dt |> 
  left_join(state_geo |> sf::st_drop_geometry() |> select(iso3,iso2, loc_id),
            join_by(country_code==loc_id)) |> 
  filter(
    !is.na(iso3),
    year>=2005
  )

pop_age_wpp_C <- popAge5dt |> 
  left_join(state_geo |> sf::st_drop_geometry() |> select(iso3,iso2, loc_id),
            join_by(country_code==loc_id)) |> 
  filter(
    !is.na(iso3),
    year %in% c(2005,2020)
  ) |> 
  filter(iso3 == iso3_SUR) |> 
  # filter(year==max(year)) |> 
  select(name, year, age, popM, popF) |> 
  rename(Male=popM, Female=popF) |> 
  mutate(age = fct_inorder(age)) |> 
  pivot_longer(cols = Male:Female,
               names_to = "Gender",
               values_to = "pop"
  ) |> 
  mutate(Gender=fct(Gender)) |> 
  group_by(year) |> 
  mutate(prop = pop/sum(pop, na.rm = TRUE)) |> 
  ungroup()

fertility_rate_C <- tfr1dt |> 
  left_join(state_geo |> sf::st_drop_geometry() |> select(iso3,iso2, loc_id),
            join_by(country_code==loc_id)) |> 
  filter(
    !is.na(iso3),
    year>=2005
  ) |> 
  filter(iso3 == iso3_SUR) 

relabel_na <- function(x) {
  x <- as.character(x)
  x[is.na(x)] <- "No data"
  factor(x)
}
```

```{r}
UHC_estimate <- UHC_all |>
  filter(IndicatorCode == "UHC_INDEX_REPORTED",
         country_name == params$country_name,
         YEAR == max(YEAR)
  )

UHC_RMNCH_estimate <- UHC_all |>
  filter(IndicatorCode == "UHC_SCI_RMNCH",
         country_name == params$country_name,
         YEAR == max(YEAR)
  )

antenatal_WHO_C <- antenatal_WHO |> filter(country_iso_3_code == iso3_SUR)
antenatal_WHO_text <- if(is_empty(antenatal_WHO_C |> 
                                  filter(year == max(year)) |> 
                                  pull(antenatal_WHO))){
  "Data not available"
} else{
  paste0(antenatal_WHO_C |> 
           filter(year == max(year)) |> 
           pull(antenatal_WHO), 
         "% in ", 
         antenatal_WHO_C |> 
           filter(year == max(year)) |> 
           pull(year)
  )
}

skilled_birth_C <- skilled_birth |> filter(COUNTRY == iso3_SUR)
skilled_birth_text <- if(is_empty(skilled_birth_C |> 
                                  filter(YEAR == max(YEAR)) |> 
                                  pull(NumericValue))){
  "Data not available"
} else{
  paste0(skilled_birth_C |> 
           filter(YEAR == max(YEAR)) |> 
           pull(NumericValue), 
         "% in ", 
         skilled_birth_C |> 
           filter(YEAR == max(YEAR)) |> 
           pull(YEAR)
  )
}


#########

life_expectancy_C <- life_expectancy |> filter(iso3 == iso3_SUR)
life_expectancy_text <- paste0(life_expectancy_C |> 
                                 filter(year==max(year)) |>  
                                 pull(e0B) |> round(1), 
                               " years in ",
                               life_expectancy_C |> 
                                 filter(year==max(year)) |> 
                                 pull(year),
                               " (",
                               life_expectancy_C |> 
                                 filter(year==min(year)) |>  
                                 pull(e0B) |> round(1),
                               " in ",
                               life_expectancy_C |> 
                                 filter(year==min(year)) |> 
                                 pull(year),
                               ")")

pop_C <- pop_wpp |> 
  filter(iso3==iso3_SUR)
pop_text <- paste0(round(pop_C |> filter(year==max(year)) |> pull(pop)/1000,1),
                   " million in ",
                   pop_C |> filter(year==max(year)) |> pull(year),
                   " (",
                   sprintf("%+.0f", (round(pop_C |> filter(year==max(year)) |> pull(pop)/pop_C |> filter(year==min(year)) |> pull(pop),2)-1)*100),
                   "% since ",
                   pop_C |> filter(year==min(year)) |> pull(year),
                   ")")

fertility_rate_text <- paste0(fertility_rate_C |> 
                                filter(year==max(year)) |>  
                                pull(tfr) |> round(1), 
                              " births per woman in ",
                              fertility_rate_C |> 
                                filter(year==max(year)) |> 
                                pull(year),
                              " (",
                              fertility_rate_C |> 
                                filter(year==min(year)) |>  
                                pull(tfr) |> round(1),
                              " in ",
                              fertility_rate_C |> 
                                filter(year==min(year)) |> 
                                pull(year),
                              ")")

constitutions_C <- constitutions |> filter(iso3 == iso3_SUR)


```

```{r, warning=FALSE}
#####################
HPV_coverage_C <- HPV_coverage |> filter(SpatialDim==iso3_SUR)
HPV_coverage_text <- if(is_empty(HPV_coverage_C |> filter(YEAR == max(YEAR)) |> pull(NumericValue))){
  "Data not available"
} else{
  paste0(HPV_coverage_C |> filter(YEAR == max(YEAR)) |> pull(NumericValue), 
         "% in ", HPV_coverage_C |> filter(YEAR == max(YEAR)) |> pull(YEAR))
}

family_planning_C <- family_planning |> filter(SpatialDim==iso3_SUR)
family_planning_text <- if(is_empty(family_planning_C |> filter(YEAR == max(YEAR)) |> pull(NumericValue))){
  "Data not available"
} else{
  paste0(family_planning_C |> filter(YEAR == max(YEAR)) |> pull(Value), 
         "% in ", family_planning_C |> filter(YEAR == max(YEAR)) |> pull(YEAR))
}

unintended_pregnancy_C <- unintended_pregnancy |> 
  filter(SpatialDim==iso3_SUR, 
         Dim1 == "UNCERTAINTY_INTERVAL_UI95")
unintended_pregnancy_text <- if(is_empty(unintended_pregnancy_C |> filter(YEAR == max(YEAR)) |> pull(NumericValue))){
  "Data not available"
} else{
  paste0(unintended_pregnancy_C |> filter(YEAR == max(YEAR)) |> pull(NumericValue) |> round(0), 
         " per 1,000 women in ", unintended_pregnancy_C |> filter(YEAR == max(YEAR)) |> pull(YEAR))
}


HPV_national_C <- HPV_national |> filter(SpatialDim==iso3_SUR)
HPV_national_text <- if(is_empty(HPV_national_C |> filter(YEAR == max(YEAR)) |> pull(Value))){
  "Data not available"
} else{
  paste0(HPV_national_C |> filter(YEAR == max(YEAR)) |> pull(Value), 
         " (evaluated in ", HPV_national_C |> filter(YEAR == max(YEAR)) |> pull(YEAR), ")")
}

HPV_national_alt_C <- HPV_national_alt |> filter(ISO_3_CODE==iso3_SUR)
HPV_national_alt_text <- if(is_empty(HPV_national_alt_C |> pull(INTRO_text))){
  "Data not available"
} else{
  paste0(HPV_national_alt_C |> pull(INTRO_text))
}

#####################
DTP3_coverage_C <- DTP3_coverage |> filter(SpatialDim==iso3_SUR)
DTP3_coverage_text <- if(is_empty(DTP3_coverage_C |> filter(YEAR == max(YEAR)) |> pull(NumericValue))){
  "Data not available"
} else{
  paste0(DTP3_coverage_C |> filter(YEAR == max(YEAR)) |> pull(NumericValue), 
         "% in ", DTP3_coverage_C |> filter(YEAR == max(YEAR)) |> pull(YEAR))
}

#####################
nursing_density_C <- nursing_density |> filter(SpatialDim==iso3_SUR)
nursing_density_text <- if(is_empty(nursing_density_C |> filter(YEAR == max(YEAR)) |> pull(NumericValue))){
  "Data not available"
} else{
  paste0(round(nursing_density_C |> filter(YEAR == max(YEAR)) |> pull(NumericValue),0), 
         " in ", nursing_density_C |> filter(YEAR == max(YEAR)) |> pull(YEAR))
}

####################
adolescent_birth_rate_C <- adolescent_birth_rate |> filter(SpatialDim==iso3_SUR)
adolescent_birth_rate_text_10_14 <- if(is_empty(adolescent_birth_rate_C |> 
                                                filter(Dim2=="AGEGROUP_YEARS10-14") |> 
                                                filter(YEAR == max(YEAR)) |> 
                                                pull(NumericValue))){
  "Data not available"
} else{
  paste0(adolescent_birth_rate_C |> 
           filter(Dim2=="AGEGROUP_YEARS10-14") |> 
           filter(YEAR == max(YEAR)) |> 
           pull(Value), 
         " per 1,000 women in ", 
         adolescent_birth_rate_C |> 
           filter(Dim2=="AGEGROUP_YEARS10-14") |> 
           filter(YEAR == max(YEAR)) |> 
           pull(YEAR)
  )
}
adolescent_birth_rate_text_15_19 <- if(is_empty(adolescent_birth_rate_C |> 
                                                filter(Dim2=="AGEGROUP_YEARS15-19") |> 
                                                filter(YEAR == max(YEAR)) |> 
                                                pull(NumericValue))){
  "Data not available"
} else{
  paste0(adolescent_birth_rate_C |> 
           filter(Dim2=="AGEGROUP_YEARS15-19") |> 
           filter(YEAR == max(YEAR)) |> 
           pull(Value), 
         " per 1,000 women in ", 
         adolescent_birth_rate_C |> 
           filter(Dim2=="AGEGROUP_YEARS15-19") |> 
           filter(YEAR == max(YEAR)) |> 
           pull(YEAR)
  )
}

##############
contraceptive_prevalence_C <- contraceptive_prevalence |> 
  filter(Dim1Type == "AGEGROUP") |> 
  filter(SpatialDim==iso3_SUR)

contraceptive_prevalence_text_15_19 <- 
  if(is_empty(contraceptive_prevalence_C |> 
              filter(Dim1=="AGEGROUP_YEARS15-19") |> 
              filter(YEAR == max(YEAR)) |> 
              pull(NumericValue))){
    "Data not available"
  } else{
    paste0(round(contraceptive_prevalence_C |> 
                   filter(Dim1=="AGEGROUP_YEARS15-19") |> 
                   filter(YEAR == max(YEAR)) |> 
                   pull(NumericValue),1), 
           "% in ", 
           contraceptive_prevalence_C |> 
             filter(Dim1=="AGEGROUP_YEARS15-19") |> 
             filter(YEAR == max(YEAR)) |> 
             pull(YEAR)
    )
  }
contraceptive_prevalence_text_20_49 <- 
  if(is_empty(contraceptive_prevalence_C |> 
              filter(Dim1=="AGEGROUP_YEARS20-49") |> 
              filter(YEAR == max(YEAR)) |> 
              pull(NumericValue))){
    "Data not available"
  } else{
    paste0(round(contraceptive_prevalence_C |> 
                   filter(Dim1=="AGEGROUP_YEARS20-49") |> 
                   filter(YEAR == max(YEAR)) |> 
                   pull(NumericValue),1),
           "% in ", 
           contraceptive_prevalence_C |> 
             filter(Dim1=="AGEGROUP_YEARS20-49") |> 
             filter(YEAR == max(YEAR)) |> 
             pull(YEAR)
    )
  }

```
\maketitle
# \underline{`r params$country_name`}
<!-- Some brief text here, for context, that runs the \textcolor{CeHDIRed}{entire width of the page}. Indicator results, tables, and figures can go into the columns below. -->
```{=latex}
\begin{multicols}{2}
\raggedcolumns     %new code
\interlinepenalty=10000
```
\begin{center}\textcolor{fgcolor}{\textbf{\hlCeHDIRed{\ \ \ \ The Right to Health \ \ \ \ }}}\end{center}  
As enshrined in Article 12 of the International Convenant on Economic, Social and Cultural Rights, the right to health is an inclusive human right that extends beyond timely and appropriate health care to encompass the underlying determinants of health.  

Under the Right to Health, States have the following obligations:  
\ \ - \textbf{{Respect:}} refrain from directly or indirectly interfering with the enjoyment of the right to health.  
\ \ - \textbf{{Protect:}} take effective measures to prevent third parties from undermining of violating the guarantees of the right to health.  
\ \ - \textbf{{Fulfill:}} adopt appropriate legislative, administrative, budgetary, judicial, promotional, and other measures toward the full realization of the right to health.
\begin{center}\textcolor{fgcolor}{\textbf{\hlCeHDIRed{\ \ \ \ The Universal Periodic Review (UPR) \ \ \ \ }}}\end{center}  
The UPR is a **State-led mechanism** that aims to evaluate each State’s human rights obligations and commitments. The review process is participatory and includes interactive discussions during which any UN Member State may issue recommendations to the State under review, which may then choose to **"Support"** or **"Note"** those recommendations.  

A CeHDI analysis of recommendations from all cycles of the UPR was used to identify health-related recommendations, which have globally gained prominence in the UPR over time (for `r params$country_name`, health-related recommendations accounted for `r health_rec_C |> filter(cycle == "Cycle 1", health_related == "Health-related") |> pull(perc)` of all recommendations in Cycle 1, and `r health_rec_C |> filter(health_related == "Health-related") |> slice_max(order_by = cycle) |> pull(perc)` in `r health_rec_C |> filter(health_related == "Health-related") |> slice_max(order_by = cycle) |> pull(cycle)`). These recommendations were further categorized into the following themes:\footnote{Visit \href{https://cehdi-har.share.connect.posit.cloud/}{cehdi-har.share.connect.posit.cloud} for more details.\label{fn:cehdiHAR}}
```{r, fig.width=5.2, fig.height=4}
a_1 <- 
  sdg_data |>    
  filter(state_under_review == params$country_name) |> 
  select(cycle, state_under_review, health_related:other_health_related, response_upr) |>
  group_by(response_upr) |>
  summarise(across(c(health_related:other_health_related), ~ sum(.x != "Other"))) |>
  ungroup() |>
  filter(response_upr %in% c("Supported", "Noted")) |>
  pivot_longer(
    cols = health_related:other_health_related,
    names_to = "theme",
    values_to = "n"
  )

a_2 <- 
  sdg_data |>    
  filter(state_under_review == params$country_name) |> 
  select(cycle, state_under_review, health_related:other_health_related, response_upr) |>
  group_by(response_upr) |>
  summarise(across(c(health_related:other_health_related), ~ sum(.x == "Other"))) |>
  ungroup() |>
  filter(response_upr %in% c("Supported", "Noted")) |>
  pivot_longer(
    cols = health_related:other_health_related,
    names_to = "theme",
    values_to = "n_other"
  )

a <- left_join(a_1, a_2) |>
  group_by(theme) |>
  mutate(
    n_tot = sum(n) + sum(n_other),
    n_tot_theme = sum(n)
  ) |>
  mutate(
    perc = n / n_tot * 100,
    perc_theme = n_tot_theme / n_tot * 100
  ) |>
  group_by(theme) |>
  mutate(
    n_sup = paste0("(", sprintf("%1.0f", n / sum(n) * 100), "%)"),
    n_sup = case_when(n_tot_theme == 0 ~ "", .default = n_sup)
  ) |>
  ungroup() |>
  filter(!theme %in% c("health_related", "TB_malaria", "NTD")) |>
  left_join(theme_labels, by = c("theme" = "variable")) |>
  arrange(-n_tot_theme) |>
  mutate(
    theme_label = case_when(is.na(theme_label) ~ theme, .default = theme_label),
    theme_label = fct_inorder(theme_label)
  )


max_a <- max(a$perc_theme, na.rm = TRUE)
a |>
  ggplot(aes(x = perc, y = fct_rev(theme_label))) +
  geom_col(aes(fill = response_upr), alpha = 0.8, width = 0.85) +
  scale_fill_manual(values = c("#ec5557", "#1c164d"))+
  labs(
    x = "Proportion of all recommendations (%)", y = NULL,
    fill = "State's response",
    title = paste("All health-related recommendations of the UPR\n", params$country_name),
    caption = "*Numbers after the bars indicate N (% supported)"
  ) +
  theme_classic() +
  scale_x_continuous(
    labels = function(x) paste0(x, "%"),
    # limits = c(0, max_a + 2),
    expand = expansion(mult = c(0, 0.15))
  ) +
  coord_cartesian(clip = "off")+
  theme(
    plot.margin = margin(l=2,t=2,b=2, r = 30, unit = "pt"),
    legend.position = c(0.99, 0.01),
    legend.justification = c("right", "bottom"),
    legend.margin = margin(0,0,0,0),
    legend.frame = element_blank(),
    legend.text = element_text(size = 9, color = "#1c164d"),
    legend.title = element_text(size = 11, color = "#1c164d"),
    legend.background = element_blank(),
    legend.key.size = unit(10,"pt"),
    axis.text.y = element_text(size = 9, color = "#1c164d"),
    axis.text.x = element_text(size = 10, color = "#1c164d"),
    plot.title = element_text(hjust = 0.5, face = "bold", color = "#1c164d"),
    axis.title.y = element_blank(),
    axis.title.x = element_text(color = "#1c164d"),
    plot.title.position = "plot",
    panel.grid = element_blank(),
    plot.caption = element_text(color = "#1c164d"),
    plot.background = element_rect(color = "#1c164d", fill = NA),
    panel.background = element_blank()
  ) +
  geom_text(
    data = a |> filter(response_upr == "Supported"),
    aes(label = paste0(n_tot_theme, " ", n_sup), x = perc_theme),
    hjust = -0.15, size = 3, vjust = 0.25, color = "#1c164d"
  )


# print(params$upr_all+
#         labs(title = params$country_name)+
#         theme(
#           axis.title.x = element_text(color = "#1c164d"),
#           axis.text = element_text(color = "#1c164d"),
#           plot.title = element_text(color = "#1c164d"),
#           plot.caption = element_text(color = "#1c164d"),
#           plot.background = element_rect(color = "#1c164d", fill = NA),
#           panel.background = element_blank(),
#           legend.background = element_blank()
#         )
# )
```
\begin{center}\textcolor{fgcolor}{\textbf{\hlCeHDIRed{\ \ \ \ Legal Basis for Health in `r params$country_name` \ \ \ \ }}}\end{center}  
\textbf{{Constitutional guarantees for citizens (as of June 2024):}}\footnote{\href{https://www.worldpolicycenter.org/maps-data/data-download/constitutions-data-download}{worldpolicycenter.org} \label{fn:worldpolicy}}  
\ \ - Right to health: `r constitutions |> filter(iso3 == iso3_SUR) |> pull(const_health)`  
\ \ - Right to public health: `r constitutions |> filter(iso3 == iso3_SUR) |> pull(const_publichealth)`  
\ \ - Right to medical care: `r constitutions |> filter(iso3 == iso3_SUR) |> pull(const_medcare)`  
\ \ - Right to a healthy environment: `r constitutions |> filter(iso3 == iso3_SUR) |> pull(const_enviro)`  
\textbf{{Legal status of abortion (as of June 2023):}}\footnote{\href{https://reproductiverights.org/wp-content/uploads/2023/12/WALM_2023-v3-Updated_12-20-23.pdf}{reproductiverights.org} \label{fn:reproductiverights}} `r world_abortion_laws |> filter(country == params$country_name) |> pull(category)`
\columnbreak
\begin{center}\textcolor{fgcolor}{\textbf{\hlCeHDIRed{\ \ \ \ Demographics \ \ \ \ }}}\end{center}  
\textbf{{Population:}}\footnote{\href{https://population.un.org/wpp}{population.un.org/wpp} \label{fn:WPP}} `r pop_text`  
\textbf{{Life expectancy at birth:}}\textsuperscript{\ref{fn:WPP}} `r life_expectancy_text`
\begin{center}\textcolor{fgcolor}{\textbf{\hlCeHDIRed{\ \ \ \ Sexual and Reproductive Health \ \ \ \ }}}\end{center}  
\textbf{{Total fertility rate:}}\textsuperscript{\ref{fn:WPP}} `r fertility_rate_text`  
\textbf{{Met need for family planning:}}\textsuperscript{\ref{fn:whoGHO}} `r family_planning_text`  
\textbf{{Adolescent birth rate:}}\footnote{\href{https://www.who.int/data/gho/info/gho-odata-api}{WHO Global Health Observatory} \label{fn:whoGHO}}  
\ \ - Aged 10 to 14: `r adolescent_birth_rate_text_10_14`  
\ \ - Aged 15 to 19: `r adolescent_birth_rate_text_15_19`  
\textbf{{Births attended by skilled health personnel:}}\textsuperscript{\ref{fn:whoGHO}} `r skilled_birth_text`  
\textbf{{Maternal mortality ratio (MMR):}}\textsuperscript{\ref{fn:whoGHO}} `r MMR |> filter(SpatialDim == iso3_SUR, YEAR == 2023) |> pull(NumericValue) |> round(digits = 0)` maternal deaths per 100,000 live births in 2023 (`r MMR |> filter(SpatialDim == iso3_SUR, YEAR == 2005) |> pull(NumericValue) |> round(digits = 0)` in 2005)  
```{r, out.width="90%"}
# mmr_estimate_2023 = params$mmr_map_plot@data |>
#   filter(country_name == params$country_name,
#          YEAR == "2023") |>
#   pull(NumericValue) |>
#   round(0)
# 
# map_insetting(
#   p1 = params$mmr_map_plot,
#   p_caption_text = paste0(params$country_name, ": ", mmr_estimate_2023, " per 100,000 live births"),
#   legend_title_text = "MMR",
#   p_title_text = "Maternal mortality ratio (MMR) estimates in 2023",
#   bbox_SUR_region_dynamic = params$bbox_SUR_region_dynamic,
#   bbox_sur = params$bbox_selected_SUR,
#   sur_area = params$sur_area
# )+labs(fill="MMR")

p1<-MMR |> 
  filter(YEAR==2023, !is.na(country_name)) |> 
  right_join(state_geo, by=join_by(SpatialDim==iso3)) |> 
  mutate(selected_sur = factor(case_when(
    country_name == params$country_name ~ params$country_name,
    .default = "Other"
  ),
  levels = c(params$country_name, "Other")
  )) |> 
  ggplot(aes(geometry = polygon))+
  geom_sf(aes(fill=mmr_cat), color = NA)+
  scale_fill_brewer(palette = "YlOrRd", na.value = "grey80", labels = relabel_na)+
  theme_void()+
  labs(fill="MMR\nin 2023")

p2 <- p1 +
  geom_sf(color="grey70", fill=NA)+
  geom_rect(
    aes(
      xmin = params$bbox_selected_SUR["xmin"] - 0.7,
      xmax = params$bbox_selected_SUR["xmax"] + 0.7,
      ymin = params$bbox_selected_SUR["ymin"] - 0.7,
      ymax = params$bbox_selected_SUR["ymax"] + 0.7
    ),
    fill = "transparent",
    color = "red",
    linewidth = 0.5
  )+
  coord_sf(
    # expand = FALSE,
    xlim = c(params$bbox_SUR_region_dynamic[[1]],params$bbox_SUR_region_dynamic[[3]]),
    ylim = c(params$bbox_SUR_region_dynamic[[2]], params$bbox_SUR_region_dynamic[[4]])
  )

p3<-p1+
  guides(fill="none")+
  ggnewscale::new_scale_fill()+ 
  geom_sf(
    aes(fill = selected_sur),
    color = "transparent")+ 
  scale_fill_manual(values=c("transparent", "white"))+
  guides(fill="none")+
  coord_sf(
    expand=FALSE,
    xlim = c(params$bbox_selected_SUR[[1]]-.11, params$bbox_selected_SUR[[3]]+.11),
    ylim = c(params$bbox_selected_SUR[[2]]-.11, params$bbox_selected_SUR[[4]]+.11)
  )+
  theme(panel.background = element_rect(fill="white"))+
  geom_rect(
    aes(
      xmin = params$bbox_selected_SUR["xmin"]-.1,
      xmax = params$bbox_selected_SUR["xmax"]+.1,
      ymin = params$bbox_selected_SUR["ymin"]-.1,
      ymax = params$bbox_selected_SUR["ymax"]+.1
    ),
    fill = "transparent",
    color = "red",
    linewidth = 0.5
  )

a<-ggplot_build(p2)
b<-ggplot_build(p3)
multiplier_a <- abs(a$layout$coord$limits$x[1]-a$layout$coord$limits$x[2])/abs(a$layout$coord$limits$y[1]-a$layout$coord$limits$y[2])
multiplier_b <- abs(b$layout$coord$limits$x[1]-b$layout$coord$limits$x[2])/abs(b$layout$coord$limits$y[1]-b$layout$coord$limits$y[2])

area_a <- abs(a$layout$coord$limits$x[1]-a$layout$coord$limits$x[2])*abs(a$layout$coord$limits$y[1]-a$layout$coord$limits$y[2])
area_b <- abs(b$layout$coord$limits$x[1]-b$layout$coord$limits$x[2])*abs(b$layout$coord$limits$y[1]-b$layout$coord$limits$y[2])

key_dim <- 0.3/(multiplier_a/multiplier_b)
inset_dimensions <- if(multiplier_b<=3){c(0,0,key_dim,key_dim*multiplier_a/multiplier_b)}else{c(0,0,key_dim*multiplier_a/multiplier_b,0.2)}

if (
  # sur_area > 11^11 | 
  area_b > 0.02*area_a
) {
  p2
} else {
  p2 + inset_element(p3, inset_dimensions[1], inset_dimensions[2],inset_dimensions[3],inset_dimensions[4])+ plot_annotation(
    theme = theme(plot.margin = margin(0,0,0,0), plot.background = element_blank()
    ))
}

# params$mmr_map_plot

# params$mmr_map_plot+theme(
#   plot.background = element_blank()
# )&
# plot_annotation(theme = theme(
#   plot.background = element_rect(color = "#1c164d", fill=NA)
#   )
#   )
```
```{r, fig.width=8, fig.height=1.5}
start_year <- "2005"
dat_plot <- MMR |>
  mutate(selected_sur = factor(case_when(
    country_name == params$country_name ~ params$country_name,
    .default = "Other"
  ),
  levels = c(params$country_name, "Other")
  )) |>
  filter(country_name %in% c(
    params$country_name,
    nearest_neighbors_list[, params$country_name][1:5]
  )) |>
  filter(year >= ymd(paste0(start_year, "-01-01"))) |>
  mutate(country_name = fct_relevel(country_name, params$country_name)) |>
  group_by(country_name) |>
  mutate(
    num_stand = NumericValue - NumericValue[YEAR == start_year],
    num_low = Low - NumericValue[YEAR == start_year],
    num_high = High - NumericValue[YEAR == start_year]
  ) |>
  ungroup()

hline_data <- dat_plot |>
  filter(YEAR == as.numeric(start_year))
str_max <- max(str_width(dat_plot |> pull(country_name)))
str_max_height <- if(str_max<21){0.8} else if(str_max <37){1.5} else{2.1}

dat_plot |>
  mutate(country_name = fct_relevel(country_name, params$country_name)) |>
  ggplot(aes(x = YEAR, y = NumericValue)) +
  labs(
    # title = paste0("Trends in MMR, neighboring States plotted for comparison"),
    x = NULL, y = "MMR",
    color = NULL,
    fill = NULL
  ) +
  scale_x_continuous(breaks = c(2005, 2020), 
                     expand = expansion(mult = c(0, 0)))+
  geom_line(lwd = 1, aes(color = selected_sur)) +
  geom_ribbon(aes(ymin = Low, ymax = High, fill = selected_sur), color = NA, alpha = 0.4) +
  scale_color_manual(values = c("#ec5557", "#1c164d")) +
  scale_fill_manual(values = c("#ec5557", "#1c164d")) +
  guides(color = "none", lwd = "none", fill = "none") +
  facet_wrap(. ~ country_name, nrow=1, scales = "free"
             # ,labeller = labeller(country_name = label_wrap_gen(21))
  ) +
  geom_hline(data = hline_data, aes(yintercept = NumericValue), lty = 2) +
  theme_bw()+
  theme(
    plot.margin = margin(0,0,0,0),
    axis.text = element_text(color = "#1c164d"),
    axis.title = element_text(color = "#1c164d"),
    axis.line = element_line(color = "#1c164d"),
    axis.ticks = element_line(color = "#1c164d"),
    panel.border = element_blank(),
    margins = margin(0,0,0,0),
    plot.background = element_blank(),
    panel.background = element_blank(),
    strip.background =element_rect(fill="#1c164d", color = "#1c164d"),
    # strip.text = element_text(colour = 'white'),
    strip.text = element_textbox_simple(
      color="white",
      height = unit(str_max_height, "lines"),
      hjust = 0.5, 
      vjust = 0.5, 
      halign = 0.5,
      valign = 0.5
    )
  )+
  coord_cartesian(clip="off")
```
<!-- \textbf{{Contraceptive prevalence (modern methods):}}\textsuperscript{\ref{fn:whoGHO}}   -->
<!-- \ \ - Aged 15 to 19:  `r contraceptive_prevalence_text_15_19`   -->
<!-- \ \ - Aged 20 to 49:  `r contraceptive_prevalence_text_20_49` -->
\begin{center}\textcolor{fgcolor}{\textbf{\hlCeHDIRed{\ \ \ \ Health Services Coverage (latest available data) \ \ \ \ }}}\end{center}  
<!-- \textbf{{UHC service coverage index:}}\textsuperscript{\ref{fn:whoGHO}} `r UHC_estimate |> pull(Value)` in `r UHC_estimate |> pull(YEAR)`   -->
\textbf{{UHC service coverage (`r UHC_all |> filter(COUNTRY == iso3_SUR, IndicatorCode == "UHC_INDEX_REPORTED", YEAR == min(YEAR)) |> pull(YEAR)` to `r UHC_all |> filter(COUNTRY == iso3_SUR, IndicatorCode == "UHC_INDEX_REPORTED", YEAR == max(YEAR)) |> pull(YEAR)`):}}\textsuperscript{\ref{fn:whoGHO}}
```{r, fig.align="center", fig.width=7, fig.height=1.5, out.width="90%"}
uhc_plots <- function(dat, title_text = "Update title", yMAX = 100){
  p_data <- dat |> 
    filter(COUNTRY == iso3_SUR)
  label_data_1 <- p_data |> 
    filter(YEAR %in% c(min(YEAR)))
  label_data_2 <- p_data |> 
    filter(YEAR %in% c(max(YEAR)))
  p1<-p_data |> 
    # mutate(val_maxtime = NumericValue[TimeDim==max(TimeDim)])
    ggplot(aes(x=YEAR, y= NumericValue))+
    geom_hline(aes(yintercept = yMAX), color = "green", lty=2, lwd=1)+
    geom_line(color = "#ec5557", lwd=1, fill=NA)+
    geom_text(
      data = label_data_1, # Use the special, single-row data frame
      aes(label = Value, # The text for the label
          nudge_y = if_else(label_data_1$NumericValue> 30, -12,13)
      ),      # Nudge the label slightly
      size=5, hjust = 0,
      color = "#1c164d", # Match theme color
      # fontface = "bold" # Make it stand out
    )+
    geom_text(
      data = label_data_2, # Use the special, single-row data frame
      aes(label = Value, # The text for the label
          nudge_y = if_else(label_data_2$NumericValue> 30, -12,13)
      ),      # Nudge the label slightly
      size=5, hjust = 1,
      color = "#1c164d", # Match theme color
      # fontface = "bold" # Make it stand out
    )+
    scale_y_continuous(limits = c(0,yMAX))+
    theme_void()+
    labs(title = title_text)+
    theme(
      axis.line = element_line(color = "#1c164d"),
      plot.title = element_text(color = "#1c164d", size = 10, hjust = 0.5))+
    coord_cartesian(clip = "off")
  p1
}

p1<-uhc_plots(UHC_all |> filter(IndicatorCode == "UHC_INDEX_REPORTED"), 
              title_text = "Services Coverage\nIndex (max: 80)", yMAX = 80)
p2<-uhc_plots(UHC_all |> filter(IndicatorCode == "UHC_SCI_CAPACITY"), 
              title_text = "Sub-index on service\ncapacity and access")
p3<-uhc_plots(UHC_all |> filter(IndicatorCode == "UHC_SCI_INFECT"), 
              title_text = "Sub-index on\ninfectious diseases")
p4<-uhc_plots(UHC_all |> filter(IndicatorCode == "UHC_SCI_NCD"), 
              title_text = "Sub-index on NCDs")
p5<-uhc_plots(UHC_all |> filter(IndicatorCode == "UHC_SCI_RMNCH"), 
              title_text = "Sub-index on RMNCH")

(p1 | p2 | p3 | p4 | p5)+plot_annotation(
  theme = theme(plot.margin = margin(0,0,0,0), 
                plot.background = element_blank()
  ))
```
<!-- \textbf{{UHC sub-index on RMNCH:}}\textsuperscript{\ref{fn:whoGHO}} `r UHC_RMNCH_estimate |> pull(Value)` in `r UHC_RMNCH_estimate |> pull(YEAR)`   -->
\textbf{{Antenatal care coverage (\ge 4 visits):}}\footnote{\href{https://platform.who.int/data/maternal-newborn-child-adolescent-ageing}{WHO MNCAH Data Portal} \label{fn:whoMNCAH}} `r antenatal_WHO_text`  
\textbf{{Nursing/midwifery personnel (per 10,000):}}\textsuperscript{\ref{fn:whoGHO}}  `r nursing_density_text`  

\textbf{{HPV coverage (9-14 year old girls):}}\textsuperscript{\ref{fn:whoGHO}}  `r HPV_coverage_text`  
\textbf{{HPV vaccine introduction:}}\footnote{\href{https://immunizationdata.who.int}{immunizationdata.who.int} \label{fn:whoimmunization}} `r HPV_national_alt_text`  
\textbf{{DTP3 coverage (among 1-year-olds):}}\textsuperscript{\ref{fn:whoGHO}} `r DTP3_coverage_text`
```{=latex}
\end{multicols}
```
\newpage
\begin{center}\textcolor{fgcolor}{\textbf{\hlCeHDIRed{\ \ \ Recommendations from the UPR, on the themes of maternal health, family planning, and abortion \ \ \ }}}\end{center}  
```{r}
table_data <- sdg_data |> 
  filter(iso3==iso3_SUR) |> 
  filter(if_any(any_of(c("maternal_health", "abortion", "contraception")), ~ .x != "Other")) |>
  # filter(maternal_health != "Other") |> 
  select(text_2, cycle, response_upr
         # , recommending_state_upr_comma
         # , maternal_health, abortion, contraception
  ) |> 
  rename(
    Recommendation  = text_2,
    Cycle = cycle,
    Response = response_upr
    # , `Recommending State(s)` = recommending_state_upr_comma
  )

kableExtra::kbl(
  table_data, 
  format = "latex", 
  longtable = TRUE,
  booktabs = TRUE) |> 
  kableExtra::kable_styling(
    latex_options = c("striped"),
    stripe_color = "#b7b4cc", 
    font_size = 10
    # ,position = "left"
  ) |> 
  kableExtra::column_spec(1,width = "45em")
```