---
title: "Maternal Health and the UPR"
params:
  country_name: "Switzerland"
  upr_all: NULL
  rec_plot: NULL
  mmr_map_plot: NULL
  bbox_SUR_region_dynamic: NULL
  bbox_selected_SUR: NULL
  sur_area: NULL
documentclass: article
# classoption: twocolumn
output: 
  pdf_document:
    includes:
      in_header: preamble.tex
    keep_tex: true
    latex_engine: xelatex
geometry: 
  - top=2cm
  - bottom=0.5cm
  - left=0.2cm
  - right=0.2cm
papersize: a4
date: "2025-09-30"
---
```{r setup, include=FALSE}
# This chunk loads the necessary packages for the report
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.align='center', out.width="100%")
pacman::p_load(
  here,
  dplyr, forcats, ggplot2, magrittr, readr, readxl, stringr, tibble, tidyr, lubridate,
  rlang,
  ggtext, 
  patchwork,
  # apyramid,
  sf
)

state_geo <- readRDS(here("output", "state_geo_enhanced.rds"))
nearest_neighbors_list <- readRDS(here("output", "nearest_neighbors_list.rds"))
iso3_SUR <- state_geo |> filter(country==params$country_name) |> pull(iso3)

# Loop through API-generated files
for (file_name in list.files(path = here("data", "API_data"), pattern = "\\.rds$")) {
  object_name <- gsub("\\.rds$", "", file_name)
  assign(object_name, readRDS(here("data", "API_data", file_name)))
}

# Life expectancy
life_expectancy <- e01dt |> 
  left_join(state_geo |> sf::st_drop_geometry() |> select(iso3,iso2, loc_id),
            join_by(country_code==loc_id)) |> 
  filter(
    !is.na(iso3),
    year>=2005
  );rm(e01dt)

pop_wpp <- pop1dt |> 
  left_join(state_geo |> sf::st_drop_geometry() |> select(iso3,iso2, loc_id),
            join_by(country_code==loc_id)) |> 
  filter(
    !is.na(iso3),
    year>=2005
  )

pop_age_wpp_C <- popAge5dt |> 
  left_join(state_geo |> sf::st_drop_geometry() |> select(iso3,iso2, loc_id),
            join_by(country_code==loc_id)) |> 
  filter(
    !is.na(iso3),
    year %in% c(2005,2020)
  ) |> 
  filter(iso3 == iso3_SUR) |> 
  # filter(year==max(year)) |> 
  select(name, year, age, popM, popF) |> 
  rename(Male=popM, Female=popF) |> 
  mutate(age = fct_inorder(age)) |> 
  pivot_longer(cols = Male:Female,
               names_to = "Gender",
               values_to = "pop"
  ) |> 
  mutate(Gender=fct(Gender)) |> 
  group_by(year) |> 
  mutate(prop = pop/sum(pop, na.rm = TRUE)) |> 
  ungroup()

fertility_rate_C <- tfr1dt |> 
  left_join(state_geo |> sf::st_drop_geometry() |> select(iso3,iso2, loc_id),
            join_by(country_code==loc_id)) |> 
  filter(
    !is.na(iso3),
    year>=2005
  ) |> 
  filter(iso3 == iso3_SUR) 


map_insetting <- function(
    p1, 
    plot_dat,
    bbox_SUR_region_dynamic, bbox_sur, sur_area, 
    p_title_text="Update this text", 
    p_caption_text = "Update this text",
    title_size=14, title_margin=17,
    caption_size=14
) {
  p2 <- p1 + 
    geom_sf(
      # data=plot_dat, 
      color="grey80", fill="transparent")+
    coord_sf(
      # expand = FALSE,
      xlim = c(bbox_SUR_region_dynamic[[1]],bbox_SUR_region_dynamic[[3]]),
      ylim = c(bbox_SUR_region_dynamic[[2]], bbox_SUR_region_dynamic[[4]])
      
      # Alternative mapping centered on the selected State
      # xlim = c(max(-180, bbox_selected_SUR()[[1]] - 20), min(180, bbox_selected_SUR()[[3]] + 20)),
      # ylim = c(max(-55.67295, bbox_selected_SUR()[[2]] - 20), min(83.6341, bbox_selected_SUR()[[4]] + 20))
    )+
    geom_rect(
      aes(
        xmin = bbox_sur["xmin"] - 0.7,
        xmax = bbox_sur["xmax"] + 0.7,
        ymin = bbox_sur["ymin"] - 0.7,
        ymax = bbox_sur["ymax"] + 0.7
      ),
      fill = "transparent",
      color = "red",
      linewidth = 0.5
    )
  
  p3 <- p1 + guides(fill="none")+
    ggnewscale::new_scale_fill()+ 
    geom_sf(
      # data=plot_dat, 
      aes(geometry=polygon, fill = selected_sur),
      color = "transparent")+ 
    scale_fill_manual(values=c("transparent", "white"))+
    # scale_linewidth_manual(values = c(0.2, 0.1)) +
    coord_sf(
      expand=FALSE,
      xlim = c(bbox_sur[[1]]-.11, bbox_sur[[3]]+.11),
      ylim = c(bbox_sur[[2]]-.11, bbox_sur[[4]]+.11)
    ) +
    theme_void()+
    theme(
      panel.background = element_rect(fill="white"), 
      panel.border = element_rect(color = "red")
    )+
    geom_rect(
      aes(
        xmin = bbox_sur["xmin"]-.1,
        xmax = bbox_sur["xmax"]+.1,
        ymin = bbox_sur["ymin"]-.1,
        ymax = bbox_sur["ymax"]+.1
      ),
      fill = "transparent",
      color = "red",
      linewidth = 0.5
    )+
    guides(fill = "none") +
    labs(
      # subtitle = "(map zoom)", 
      caption = NULL)
  
  p_title <- plot_annotation(
    title = p_title_text,
    caption = p_caption_text,
    theme = theme(
      plot.background = element_rect(color = "#1c164d", fill=NA),
      plot.title = element_textbox_simple(
        size = title_size,
        margin = margin(t = title_margin, b = title_margin, r=0, l=0, unit = "pt")
      ),
      plot.caption = element_textbox_simple(
        size = caption_size, halign = 0.5,
        margin = margin(t = 10, b = 2, r=0, l=0, unit = "pt")
      )
    )
  )
  
  # Calculate Inset dimensions for lower left
  a<-ggplot_build(p2)
  b<-ggplot_build(p3)
  multiplier_a <- abs(a$layout$coord$limits$x[1]-a$layout$coord$limits$x[2])/abs(a$layout$coord$limits$y[1]-a$layout$coord$limits$y[2])
  multiplier_b <- abs(b$layout$coord$limits$x[1]-b$layout$coord$limits$x[2])/abs(b$layout$coord$limits$y[1]-b$layout$coord$limits$y[2])
  
  area_a <- abs(a$layout$coord$limits$x[1]-a$layout$coord$limits$x[2])*abs(a$layout$coord$limits$y[1]-a$layout$coord$limits$y[2])
  area_b <- abs(b$layout$coord$limits$x[1]-b$layout$coord$limits$x[2])*abs(b$layout$coord$limits$y[1]-b$layout$coord$limits$y[2])
  
  key_dim <- 0.3/(multiplier_a/multiplier_b)
  inset_dimensions <- if(multiplier_b<=3){c(0,0,key_dim,key_dim*multiplier_a/multiplier_b)}else{c(0,0,key_dim*multiplier_a/multiplier_b,0.2)}
  
  if (
    # sur_area > 11^11 | 
    area_b > 0.02*area_a
  ) {
    p2 + p_title
  } else {
    p2 + inset_element(p3, inset_dimensions[1], inset_dimensions[2],inset_dimensions[3],inset_dimensions[4]) + p_title
  }
}

```

```{r}
UHC_estimate <- UHC_all |>
  filter(IndicatorCode == "UHC_INDEX_REPORTED",
         country_name == params$country_name,
         YEAR == max(YEAR)
  )

UHC_RMNCH_estimate <- UHC_all |>
  filter(IndicatorCode == "UHC_SCI_RMNCH",
         country_name == params$country_name,
         YEAR == max(YEAR)
  )

antenatal_WHO_C <- antenatal_WHO |> filter(country_iso_3_code == iso3_SUR)
antenatal_WHO_text <- if(is_empty(antenatal_WHO_C |> 
                                  filter(year == max(year)) |> 
                                  pull(antenatal_WHO))){
  "Data not available"
} else{
  paste0(antenatal_WHO_C |> 
           filter(year == max(year)) |> 
           pull(antenatal_WHO), 
         "% in ", 
         antenatal_WHO_C |> 
           filter(year == max(year)) |> 
           pull(year)
  )
}


#########

life_expectancy_C <- life_expectancy |> filter(iso3 == iso3_SUR)
life_expectancy_text <- paste0(life_expectancy_C |> 
                                 filter(year==max(year)) |>  
                                 pull(e0B) |> round(1), 
                               " years in ",
                               life_expectancy_C |> 
                                 filter(year==max(year)) |> 
                                 pull(year),
                               " (",
                               life_expectancy_C |> 
                                 filter(year==min(year)) |>  
                                 pull(e0B) |> round(1),
                               " years in ",
                               life_expectancy_C |> 
                                 filter(year==min(year)) |> 
                                 pull(year),
                               ")")

pop_C <- pop_wpp |> 
  filter(iso3==iso3_SUR)
pop_text <- paste0(round(pop_C |> filter(year==max(year)) |> pull(pop)/1000,1),
                   " million in ",
                   pop_C |> filter(year==max(year)) |> pull(year),
                   " (",
                   sprintf("%+.0f", (round(pop_C |> filter(year==max(year)) |> pull(pop)/pop_C |> filter(year==min(year)) |> pull(pop),2)-1)*100),
                   "% since ",
                   pop_C |> filter(year==min(year)) |> pull(year),
                   ")")

fertility_rate_text <- paste0(fertility_rate_C |> 
                                filter(year==max(year)) |>  
                                pull(tfr) |> round(1), 
                              " children per woman in ",
                              fertility_rate_C |> 
                                filter(year==max(year)) |> 
                                pull(year),
                              " (",
                              fertility_rate_C |> 
                                filter(year==min(year)) |>  
                                pull(tfr) |> round(1),
                              " in ",
                              fertility_rate_C |> 
                                filter(year==min(year)) |> 
                                pull(year),
                              ")")
```

\maketitle
# \underline{`r params$country_name`}  

<!-- Some brief text here, for context, that runs the \textcolor{CeHDIRed}{entire width of the page}. Indicator results, tables, and figures can go into the columns below. -->
```{r, out.width="30%", warning=FALSE}
p1<-UHC_all |> 
  filter(country_name == params$country_name, 
         IndicatorCode == "UHC_INDEX_REPORTED") |>
  ggplot(aes(x = YEAR, y= NumericValue))+geom_line()+
  theme_void()

#####################
HPV_coverage_C <- HPV_coverage |> filter(SpatialDim==iso3_SUR)
HPV_coverage_text <- if(is_empty(HPV_coverage_C |> filter(YEAR == max(YEAR)) |> pull(NumericValue))){
  "Data not available"
} else{
  paste0(HPV_coverage_C |> filter(YEAR == max(YEAR)) |> pull(NumericValue), 
         "% in ", HPV_coverage_C |> filter(YEAR == max(YEAR)) |> pull(YEAR))
}


HPV_national_C <- HPV_national |> filter(SpatialDim==iso3_SUR)
HPV_national_text <- if(is_empty(HPV_national_C |> filter(YEAR == max(YEAR)) |> pull(Value))){
  "Data not available"
} else{
  paste0(HPV_national_C |> filter(YEAR == max(YEAR)) |> pull(Value), 
         " (evaluated in ", HPV_national_C |> filter(YEAR == max(YEAR)) |> pull(YEAR), ")")
}

HPV_national_alt_C <- HPV_national_alt |> filter(ISO_3_CODE==iso3_SUR)
HPV_national_alt_text <- if(is_empty(HPV_national_alt_C |> pull(INTRO_text))){
  "Data not available"
} else{
  paste0(HPV_national_alt_C |> pull(INTRO_text))
}

#####################
DTP3_coverage_C <- DTP3_coverage |> filter(SpatialDim==iso3_SUR)
DTP3_coverage_text <- if(is_empty(DTP3_coverage_C |> filter(YEAR == max(YEAR)) |> pull(NumericValue))){
  "Data not available"
} else{
  paste0(DTP3_coverage_C |> filter(YEAR == max(YEAR)) |> pull(NumericValue), 
         "% in ", DTP3_coverage_C |> filter(YEAR == max(YEAR)) |> pull(YEAR))
}

#####################
nursing_density_C <- nursing_density |> filter(SpatialDim==iso3_SUR)
nursing_density_text <- if(is_empty(nursing_density_C |> filter(YEAR == max(YEAR)) |> pull(NumericValue))){
  "Data not available"
} else{
  paste0(round(nursing_density_C |> filter(YEAR == max(YEAR)) |> pull(NumericValue),0), 
         " in ", nursing_density_C |> filter(YEAR == max(YEAR)) |> pull(YEAR))
}

####################
adolescent_birth_rate_C <- adolescent_birth_rate |> filter(SpatialDim==iso3_SUR)
adolescent_birth_rate_text_10_14 <- if(is_empty(adolescent_birth_rate_C |> 
                                                filter(Dim2=="AGEGROUP_YEARS10-14") |> 
                                                filter(YEAR == max(YEAR)) |> 
                                                pull(NumericValue))){
  "Data not available"
} else{
  paste0(adolescent_birth_rate_C |> 
           filter(Dim2=="AGEGROUP_YEARS10-14") |> 
           filter(YEAR == max(YEAR)) |> 
           pull(Value), 
         " per 1,000 women in ", 
         adolescent_birth_rate_C |> 
           filter(Dim2=="AGEGROUP_YEARS10-14") |> 
           filter(YEAR == max(YEAR)) |> 
           pull(YEAR)
  )
}
adolescent_birth_rate_text_15_19 <- if(is_empty(adolescent_birth_rate_C |> 
                                                filter(Dim2=="AGEGROUP_YEARS15-19") |> 
                                                filter(YEAR == max(YEAR)) |> 
                                                pull(NumericValue))){
  "Data not available"
} else{
  paste0(adolescent_birth_rate_C |> 
           filter(Dim2=="AGEGROUP_YEARS15-19") |> 
           filter(YEAR == max(YEAR)) |> 
           pull(Value), 
         " per 1,000 women in ", 
         adolescent_birth_rate_C |> 
           filter(Dim2=="AGEGROUP_YEARS15-19") |> 
           filter(YEAR == max(YEAR)) |> 
           pull(YEAR)
  )
}

##############
contraceptive_prevalence_C <- contraceptive_prevalence |> 
  filter(Dim1Type == "AGEGROUP") |> 
  filter(SpatialDim==iso3_SUR)

contraceptive_prevalence_text_15_19 <- 
  if(is_empty(contraceptive_prevalence_C |> 
              filter(Dim1=="AGEGROUP_YEARS15-19") |> 
              filter(YEAR == max(YEAR)) |> 
              pull(NumericValue))){
    "Data not available"
  } else{
    paste0(round(contraceptive_prevalence_C |> 
                   filter(Dim1=="AGEGROUP_YEARS15-19") |> 
                   filter(YEAR == max(YEAR)) |> 
                   pull(NumericValue),1), 
           "% in ", 
           contraceptive_prevalence_C |> 
             filter(Dim1=="AGEGROUP_YEARS15-19") |> 
             filter(YEAR == max(YEAR)) |> 
             pull(YEAR)
    )
  }
contraceptive_prevalence_text_20_49 <- 
  if(is_empty(contraceptive_prevalence_C |> 
              filter(Dim1=="AGEGROUP_YEARS20-49") |> 
              filter(YEAR == max(YEAR)) |> 
              pull(NumericValue))){
    "Data not available"
  } else{
    paste0(round(contraceptive_prevalence_C |> 
                   filter(Dim1=="AGEGROUP_YEARS20-49") |> 
                   filter(YEAR == max(YEAR)) |> 
                   pull(NumericValue),1),
           "% in ", 
           contraceptive_prevalence_C |> 
             filter(Dim1=="AGEGROUP_YEARS20-49") |> 
             filter(YEAR == max(YEAR)) |> 
             pull(YEAR)
    )
  }
```
```{=latex}
\begin{multicols}{2}
\raggedcolumns     %new code
\interlinepenalty=10000
```
\begin{center}\textcolor{fgcolor}{\textbf{\hlCeHDIRed{\ \ \ \ Demographics \ \ \ \ }}}\end{center}  
\textcolor{fgcolor}{\hlCeHDIBlue{Population:}}\footnote{\href{https://population.un.org/wpp}{population.un.org/wpp} \label{fn:WPP}} `r pop_text`  
```{r, out.width="90%"}
pop_age_wpp_C_split <- pop_age_wpp_C |> 
  mutate(prop = case_when(Gender == "Male" ~ -prop*100, .default = prop*100)) |>
  group_by(year, age) |> 
  mutate(midpoint_prop = case_when(year == 2005 ~ NA,
                                   year == 2020 ~ mean(prop))) |> 
  ungroup()

max_per <- max(pop_age_wpp_C_split$prop, na.rm=T)
min_per <- min(pop_age_wpp_C_split$prop, na.rm=T)
# begin ggplot
##############
ggplot()+  # default x-axis is age in years;
  
  # population data graph
  geom_col(
    data = pop_age_wpp_C_split |> filter(year == 2005),
    mapping = aes(
      x = age,
      y = prop,
      fill = Gender),
    colour = "black",                               # black color around bars
    alpha = 0.2,                                    # more transparent
    width = 1)+                                     # full width
  
  # case data graph
  geom_col(
    data = pop_age_wpp_C_split |> filter(year == 2020), 
    mapping = aes(
      x = age,                               # age categories as original X axis
      y = prop,                                # % as original Y-axis
      fill = Gender),                             # fill of bars by gender
    colour = "black",                               # black color around bars
    alpha = 1,                                      # not transparent 
    width = 0.4)+                                   # half width
  
  geom_point(
    data=pop_age_wpp_C_split,
    aes(x=age, y=midpoint_prop
        # , color=midpoint_prop>0
    ),
    color = "yellow",
    shape = "|",
    # shape = 124,
    size = 4
  )+
  
  # flip the X and Y axes to make pyramid vertical
  coord_flip()+
  
  # manually ensure that age-axis is ordered correctly
  scale_x_discrete(limits = levels(pop_age_wpp_C_split$age))+     # defined in chunk above
  
  # set percent-axis
  scale_y_continuous(
    limits = c(min_per, max_per),                                      # min and max defined above
    breaks = seq(floor(min_per), ceiling(max_per), by = 1),                # from min% to max% by 2
    labels = paste0(                                                       # for the labels, paste together...
      abs(seq(floor(min_per), ceiling(max_per), by = 1)), "%")
  )+
  
  # designate colors and legend labels manually
  scale_fill_manual(
    # values = c("#7fc97f", "#beaed4")
    values = c("#1c164d", "#ec5557")
  ) +
  # scale_color_manual(
  #   # values = c("#7fc97f", "#beaed4")
  #   values = c("#ec5557", "#1c164d")
  # ) +
  # scale_color_manual(
  #   # values = c("#7fc97f", "#beaed4")
  #   values = c("#ec5557", "#1c164d")
  # ) +
  # plot labels, titles, caption    
  labs(
    title = "Age and gender distribution in 2020, drawn over baseline in 2005",
    x = "Age category",
    y = "Percent of total",
    fill = NULL) +
  guides(color = "Midpoint")+
  # optional aesthetic themes
  theme(
    legend.position = "inside",
    legend.justification = c(1,1),
    legend.position.inside = c(1,1), 
    legend.background = element_rect(color = NA, fill = NA), 
    legend.key = element_rect(color = NA, fill=NA),
    legend.text = element_text(size=11, color = "#1c164d"),
    axis.text = element_text(size=10, color = "#1c164d"),
    axis.title = element_text(size=11, color = "#1c164d"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(color = "#1c164d", fill = NA),
    panel.background = element_rect(color = "#1c164d", fill=NA),
    axis.line = element_line(colour = "#1c164d"),
    plot.title = element_text(hjust = 0, color = "#1c164d"), 
    plot.caption = element_text(hjust=0, size=11, face = "italic"))
```
\textcolor{fgcolor}{\hlCeHDIBlue{Total fertility rate:}}\textsuperscript{\ref{fn:WPP}} `r fertility_rate_text`  
\columnbreak

\begin{center}\textcolor{fgcolor}{\textbf{\hlCeHDIRed{\ \ \ \ Key Health Indicators (latest available data) \ \ \ \ }}}\end{center}  
\textcolor{fgcolor}{\hlCeHDIBlue{Life expectancy at birth:}}\textsuperscript{\ref{fn:WPP}} `r life_expectancy_text`
<!-- Maternal Mortality Ratio:   -->
<!-- Under-Five Mortality Rate   -->  
\textcolor{fgcolor}{\hlCeHDIBlue{Adolescent birth rate:}}\footnote{\href{https://www.who.int/data/gho/info/gho-odata-api}{WHO Global Health Observatory (accessed through API)} \label{fn:whoGHO}}  
\ \ - Aged 10 to 14: `r adolescent_birth_rate_text_10_14`  
\ \ - Aged 15 to 19: `r adolescent_birth_rate_text_15_19`  

\textcolor{fgcolor}{\hlCeHDIBlue{Contraceptive prevalence (modern methods):}}\textsuperscript{\ref{fn:whoGHO}}  
\ \ - Aged 15 to 19:  `r contraceptive_prevalence_text_15_19`  
\ \ - Aged 20 to 49:  `r contraceptive_prevalence_text_20_49`
\begin{center}\textcolor{fgcolor}{\textbf{\hlCeHDIRed{\ \ \ \ Health Services Coverage (latest available data) \ \ \ \ }}}\end{center}  
\textcolor{fgcolor}{\hlCeHDIBlue{UHC service coverage index:}}\textsuperscript{\ref{fn:whoGHO}} `r UHC_estimate |> pull(Value)` in `r UHC_estimate |> pull(YEAR)`  
\textcolor{fgcolor}{\hlCeHDIBlue{UHC sub-Index on RMNCH:}}\textsuperscript{\ref{fn:whoGHO}} `r UHC_RMNCH_estimate |> pull(Value)` in `r UHC_RMNCH_estimate |> pull(YEAR)`  
\textcolor{fgcolor}{\hlCeHDIBlue{Antenatal care coverage (\ge 4 visits):}}\footnote{\href{https://platform.who.int/data/maternal-newborn-child-adolescent-ageing}{WHO MNCAH Data Portal} \label{fn:whoMNCAH}} `r antenatal_WHO_text`  
\textcolor{fgcolor}{\hlCeHDIBlue{Nursing/midwifery personnel (per 10,000):}}\textsuperscript{\ref{fn:whoGHO}}  `r nursing_density_text`  

\textcolor{fgcolor}{\hlCeHDIBlue{HPV coverage (9-14 year old girls):}}\textsuperscript{\ref{fn:whoGHO}}  `r HPV_coverage_text`  
\textcolor{fgcolor}{\hlCeHDIBlue{HPV vaccine introduction:}}\footnote{\href{https://immunizationdata.who.int}{immunizationdata.who.int} \label{fn:whoimmunization}} `r HPV_national_alt_text`  
\textcolor{fgcolor}{\hlCeHDIBlue{DTP3 coverage (among 1-year-olds):}}\textsuperscript{\ref{fn:whoGHO}} `r DTP3_coverage_text`  
```{=latex}
\end{multicols}
```
\redrule

```{=latex}
\setlength\columnseprule{.3pt}
\renewcommand{\columnseprulecolor}{\color{CeHDIRed}}
\begin{multicols}{2}
\raggedcolumns     %new code
\interlinepenalty=10000
```
<!-- \columnbreak -->

```{r, fig.width=6}
mmr_estimate_2023 = params$mmr_map_plot@data |>
  filter(country_name == params$country_name,
         YEAR == "2023") |>
  pull(NumericValue) |>
  round(0)

map_insetting(
  p1 = params$mmr_map_plot,
  p_caption_text = paste0(params$country_name, ": ", mmr_estimate_2023, " per 100,000 live births"),
  p_title_text = "Maternal mortality ratio (MMR) estimates in 2023",
  bbox_SUR_region_dynamic = params$bbox_SUR_region_dynamic,
  bbox_sur = params$bbox_selected_SUR,
  sur_area = params$sur_area
)

# params$mmr_map_plot

# params$mmr_map_plot+theme(
#   plot.background = element_blank()
# )&
# plot_annotation(theme = theme(
#   plot.background = element_rect(color = "#1c164d", fill=NA)
#   )
#   )
```

\pagebreak

\begin{center}\textcolor{fgcolor}{\textbf{\hlCeHDIRed{\ \ \ \ The Right to Health \ \ \ \ }}}\end{center}  
As enshrined in Article 12 of the International Convenant on Economic, Social and Cultural Rights, the right to health is an inclusive human right that extends beyond timely and appropriate health care to encompass the underlying determinants of health.  

It forms an essential part of States’ obligations under international human rights law and provides a binding normative framework for advancing well-being, equity, and dignity across all sectors of society.  

Under the Right to Health, States have the following obligations:
\begin{myitemize}
\item \textcolor{fgcolor}{\hlCeHDIBlue{Respect:}} refrain from directly or indirectly interfering with the enjoyment of the right to health.  
\item \textcolor{fgcolor}{\hlCeHDIBlue{Protect:}} take effective measures to prevent third parties from undermining of violating the guarantees of the right to health.  
\item \textcolor{fgcolor}{\hlCeHDIBlue{Fulfill:}} adopt appropriate legislative, administrative, budgetary, judicial, promotional, and other measures toward the full realization of the right to health.
\end{myitemize}  

\begin{center}\textcolor{fgcolor}{\textbf{\hlCeHDIRed{\ \ \ \ The Universal Periodic Review (UPR) \ \ \ \ }}}\end{center}  
The UPR is a **State-led mechanism** that aims to evaluate each State’s human rights obligations and commitments. The review process is participatory and includes interactive discussions during which any UN Member State may issue recommendations to the State under review, which may then choose to ‘support’ or ‘note’ those recommendations.  
\includegraphics[width=\columnwidth]{WHO_UPR-removebg3.png}
An analysis of recommendations from all cycles of the UPR was used to identify health-related recommendations within the following themes:\footnote{Visit \href{https://cehdi-har.share.connect.posit.cloud/}{cehdi-har.share.connect.posit.cloud} for more details.\label{fn:cehdiHAR}}
```{r, fig.width=6, fig.height=5.5}
print(params$upr_all+
        labs(title = params$country_name)+
        theme(
          axis.title.x = element_text(color = "black"),
          plot.background = element_rect(color = "#1c164d", fill = NA),
          panel.background = element_blank(),
          legend.background = element_blank()
        )
)
```

**Note:** See wrapfig package: <https://latex-tutorial.com/wrap-image-text/>
```{r, fig.width=4, fig.height=3.5, out.width="50%"}
print(
  params$rec_plot+
    labs(y="Recommendations (N)")+
    geom_text(aes(label = sprintf("%1.0f", n_tot), y = n_tot, vjust = -0.2), size = 5, fontface = "bold", color = "#1c164d") +
    theme(
      panel.grid = element_blank(),
      axis.text.x = element_text(size = 12, color = "#1c164d"),
      axis.text.y = element_text(size = 12, color = "#1c164d"),
      axis.title.x = element_blank(),
      axis.title.y = element_text(size = 14, color = "#1c164d"),
      legend.position = "bottom",
      # legend.position = c(0.01, 0.99),
      # legend.justification = c("left", "top"),
      legend.text = element_text(size = 12,colour = "#1c164d"),
      # plot.background = element_blank(),
      plot.background = element_rect(color = "#1c164d", fill = NA),
      panel.border = element_rect(color = "#1c164d"),
      # panel.border = element_blank(),
      axis.ticks = element_line(color = "#1c164d"),
      panel.background = element_blank(),
      legend.background = element_blank(),
      strip.background = element_blank(),
      strip.text = element_blank(),
      plot.title = element_blank()
    )
)
```

<!-- ```{r, fig.height=3, out.width="80%"} -->
<!-- penguins |> ggplot(aes(x = island, y = bill_len, fill=species))+geom_violin() -->
<!-- ``` -->

<!-- ```{r} -->
<!-- penguins |> ggplot(aes(x = island, y = bill_len, fill=species))+geom_boxplot() -->
<!-- ``` -->

<!-- ```{r, fig.height=3, out.width="80%"} -->
<!-- penguins |> ggplot(aes(x = island, y = bill_len, fill=species))+geom_violin() -->
<!-- ``` -->

<!-- ```{r, fig.width=3, out.height="500pt", } -->
<!-- penguins |> ggplot(aes(x = island, y = bill_len, fill=species))+geom_boxplot() -->
<!-- ``` -->

<!-- ```{r, fig.height=3, out.width="80%"} -->
<!-- penguins |> ggplot(aes(x = island, y = bill_len, fill=species))+geom_violin() -->
<!-- ``` -->

<!-- ```{r} -->
<!-- penguins |> ggplot(aes(x = island, y = bill_len, fill=species))+geom_boxplot() -->
<!-- ``` -->

<!-- ```{r, fig.height=3, out.width="80%"} -->
<!-- penguins |> ggplot(aes(x = island, y = bill_len, fill=species))+geom_violin() -->
<!-- ``` -->

<!-- ```{r} -->
<!-- penguins |> ggplot(aes(x = island, y = bill_len, fill=species))+geom_boxplot() -->
<!-- ``` -->

```{=latex}
\end{multicols}
```