---
title: "The UPR & Maternal Health"
params:
  country_name: "Switzerland"
  upr_all: NULL
  rec_plot: NULL
  mmr_map_plot: NULL
documentclass: article
# classoption: twocolumn
output: 
  pdf_document:
    includes:
      in_header: preamble.tex
    keep_tex: true
    latex_engine: xelatex
geometry: 
  - top=2cm
  - bottom=0.5cm
  - left=0.2cm
  - right=0.2cm
papersize: a4
date: "2025-09-30"
---
```{r setup, include=FALSE}
# This chunk loads the necessary packages for the report
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.align='center', out.width="100%")
pacman::p_load(
  here,
  dplyr, forcats, ggplot2, magrittr, readr, readxl, stringr, tibble, tidyr, lubridate,
  rlang,
  ggtext, 
  patchwork,
  # apyramid,
  sf
)

state_geo <- readRDS(here("output", "state_geo_enhanced.rds"))
nearest_neighbors_list <- readRDS(here("output", "nearest_neighbors_list.rds"))
iso3_SUR <- state_geo |> filter(country==params$country_name) |> pull(iso3)

# Loop through API-generated files
for (file_name in list.files(path = here("data", "API_data"), pattern = "\\.rds$")) {
  object_name <- gsub("\\.rds$", "", file_name)
  assign(object_name, readRDS(here("data", "API_data", file_name)))
}

# Life expectancy
life_expectancy <- e01dt |> 
  left_join(state_geo |> sf::st_drop_geometry() |> select(iso3,iso2, loc_id),
            join_by(country_code==loc_id)) |> 
  filter(
    !is.na(iso3),
    year>=2005
  );rm(e01dt)

pop_wpp <- pop1dt |> 
  left_join(state_geo |> sf::st_drop_geometry() |> select(iso3,iso2, loc_id),
            join_by(country_code==loc_id)) |> 
  filter(
    !is.na(iso3),
    year>=2005
  )

pop_age_wpp_C <- popAge5dt |> 
  left_join(state_geo |> sf::st_drop_geometry() |> select(iso3,iso2, loc_id),
            join_by(country_code==loc_id)) |> 
  filter(
    !is.na(iso3),
    year %in% c(2005,2020)
  ) |> 
  filter(iso3 == iso3_SUR) |> 
  # filter(year==max(year)) |> 
  select(name, year, age, popM, popF) |> 
  rename(Male=popM, Female=popF) |> 
  mutate(age = fct_inorder(age)) |> 
  pivot_longer(cols = Male:Female,
               names_to = "Gender",
               values_to = "pop"
  ) |> 
  mutate(Gender=fct(Gender)) |> 
  group_by(year) |> 
  mutate(prop = pop/sum(pop, na.rm = TRUE)) |> 
  ungroup()
```

```{r}
UHC_estimate <- UHC_all |>
  filter(IndicatorCode == "UHC_INDEX_REPORTED",
         country_name == params$country_name,
         YEAR == max(YEAR)
  )

UHC_RMNCH_estimate <- UHC_all |>
  filter(IndicatorCode == "UHC_SCI_RMNCH",
         country_name == params$country_name,
         YEAR == max(YEAR)
  )

antenatal_WHO_C <- antenatal_WHO |> filter(country_iso_3_code == iso3_SUR)
antenatal_WHO_text <- if(is_empty(antenatal_WHO_C |> 
                                                filter(year == max(year)) |> 
                                                pull(antenatal_WHO))){
  "Data not available"
} else{
  paste0(antenatal_WHO_C |> 
           filter(year == max(year)) |> 
           pull(antenatal_WHO), 
         "% in ", 
         antenatal_WHO_C |> 
           filter(year == max(year)) |> 
           pull(year)
  )
}


#########

life_expectancy_C <- life_expectancy |> filter(iso3 == iso3_SUR)
life_expectancy_text <- paste0(life_expectancy_C |> 
                                 filter(year==max(year)) |>  
                                 pull(e0B) |> round(1), 
                               " years in ",
                               life_expectancy_C |> 
                                 filter(year==max(year)) |> 
                                 pull(year),
                               " (",
                               life_expectancy_C |> 
                                 filter(year==min(year)) |>  
                                 pull(e0B) |> round(1),
                               " years in ",
                               life_expectancy_C |> 
                                 filter(year==min(year)) |> 
                                 pull(year),
                               ")")

pop_C <- pop_wpp |> 
  filter(iso3==iso3_SUR)
pop_text <- paste0(round(pop_C |> filter(year==max(year)) |> pull(pop)/1000,1),
                   " million in ",
                   pop_C |> filter(year==max(year)) |> pull(year),
                   " (",
                   sprintf("%+.0f", (round(pop_C |> filter(year==max(year)) |> pull(pop)/pop_C |> filter(year==min(year)) |> pull(pop),2)-1)*100),
                   "% since ",
                   pop_C |> filter(year==min(year)) |> pull(year),
                   ")")
```

\maketitle
# `r params$country_name`  

<!-- Some brief text here, for context, that runs the \textcolor{CeHDIRed}{entire width of the page}. Indicator results, tables, and figures can go into the columns below. -->
```{r, out.width="30%", warning=FALSE}
p1<-UHC_all |> 
  filter(country_name == params$country_name, 
         IndicatorCode == "UHC_INDEX_REPORTED") |>
  ggplot(aes(x = YEAR, y= NumericValue))+geom_line()+
  theme_void()

#####################
HPV_coverage_C <- HPV_coverage |> filter(SpatialDim==iso3_SUR)
HPV_coverage_text <- if(is_empty(HPV_coverage_C |> filter(YEAR == max(YEAR)) |> pull(NumericValue))){
  "Data not available"
} else{
  paste0(HPV_coverage_C |> filter(YEAR == max(YEAR)) |> pull(NumericValue), 
         "% in ", HPV_coverage_C |> filter(YEAR == max(YEAR)) |> pull(YEAR))
}


HPV_national_C <- HPV_national |> filter(SpatialDim==iso3_SUR)
HPV_national_text <- if(is_empty(HPV_national_C |> filter(YEAR == max(YEAR)) |> pull(Value))){
  "Data not available"
} else{
  paste0(HPV_national_C |> filter(YEAR == max(YEAR)) |> pull(Value), 
         " (evaluated in ", HPV_national_C |> filter(YEAR == max(YEAR)) |> pull(YEAR), ")")
}
####################
adolescent_birth_rate_C <- adolescent_birth_rate |> filter(SpatialDim==iso3_SUR)
adolescent_birth_rate_text_10_14 <- if(is_empty(adolescent_birth_rate_C |> 
                                                filter(Dim2=="AGEGROUP_YEARS10-14") |> 
                                                filter(YEAR == max(YEAR)) |> 
                                                pull(NumericValue))){
  "Data not available"
} else{
  paste0(adolescent_birth_rate_C |> 
           filter(Dim2=="AGEGROUP_YEARS10-14") |> 
           filter(YEAR == max(YEAR)) |> 
           pull(Value), 
         " per 1,000 women in ", 
         adolescent_birth_rate_C |> 
           filter(Dim2=="AGEGROUP_YEARS10-14") |> 
           filter(YEAR == max(YEAR)) |> 
           pull(YEAR)
  )
}
adolescent_birth_rate_text_15_19 <- if(is_empty(adolescent_birth_rate_C |> 
                                                filter(Dim2=="AGEGROUP_YEARS15-19") |> 
                                                filter(YEAR == max(YEAR)) |> 
                                                pull(NumericValue))){
  "Data not available"
} else{
  paste0(adolescent_birth_rate_C |> 
           filter(Dim2=="AGEGROUP_YEARS15-19") |> 
           filter(YEAR == max(YEAR)) |> 
           pull(Value), 
         " per 1,000 women in ", 
         adolescent_birth_rate_C |> 
           filter(Dim2=="AGEGROUP_YEARS15-19") |> 
           filter(YEAR == max(YEAR)) |> 
           pull(YEAR)
  )
}

##############
contraceptive_prevalence_C <- contraceptive_prevalence |> 
  filter(Dim1Type == "AGEGROUP") |> 
  filter(SpatialDim==iso3_SUR)

contraceptive_prevalence_text_15_19 <- 
  if(is_empty(contraceptive_prevalence_C |> 
              filter(Dim1=="AGEGROUP_YEARS15-19") |> 
              filter(YEAR == max(YEAR)) |> 
              pull(NumericValue))){
    "Data not available"
  } else{
    paste0(round(contraceptive_prevalence_C |> 
             filter(Dim1=="AGEGROUP_YEARS15-19") |> 
             filter(YEAR == max(YEAR)) |> 
             pull(NumericValue),1), 
           "% in ", 
           contraceptive_prevalence_C |> 
             filter(Dim1=="AGEGROUP_YEARS15-19") |> 
             filter(YEAR == max(YEAR)) |> 
             pull(YEAR)
    )
  }
contraceptive_prevalence_text_20_49 <- 
  if(is_empty(contraceptive_prevalence_C |> 
              filter(Dim1=="AGEGROUP_YEARS20-49") |> 
              filter(YEAR == max(YEAR)) |> 
              pull(NumericValue))){
    "Data not available"
  } else{
    paste0(round(contraceptive_prevalence_C |> 
             filter(Dim1=="AGEGROUP_YEARS20-49") |> 
             filter(YEAR == max(YEAR)) |> 
             pull(NumericValue),1),
           "% in ", 
           contraceptive_prevalence_C |> 
             filter(Dim1=="AGEGROUP_YEARS20-49") |> 
             filter(YEAR == max(YEAR)) |> 
             pull(YEAR)
    )
  }
```
```{=latex}
\begin{multicols}{2}
\raggedcolumns     %new code
\interlinepenalty=10000
```
\begin{center}\textcolor{fgcolor}{\textbf{\hlCeHDIRed{Demographics}}}\end{center}  
\textcolor{fgcolor}{\hlCeHDIBlue{Population:}} `r pop_text`  
```{r, out.width="70%"}
pop_age_wpp_C_split <- pop_age_wpp_C |> 
  mutate(prop = case_when(Gender == "Male" ~ -prop*100, .default = prop*100))

max_per <- max(pop_age_wpp_C_split$prop, na.rm=T)
min_per <- min(pop_age_wpp_C_split$prop, na.rm=T)
# begin ggplot
##############
ggplot()+  # default x-axis is age in years;

  # population data graph
  geom_col(
    data = pop_age_wpp_C_split |> filter(year == 2005),
    mapping = aes(
      x = age,
      y = prop,
      fill = Gender),
    colour = "black",                               # black color around bars
    alpha = 0.2,                                    # more transparent
    width = 1)+                                     # full width
  
  # case data graph
  geom_col(
    data = pop_age_wpp_C_split |> filter(year == 2020), 
    mapping = aes(
      x = age,                               # age categories as original X axis
      y = prop,                                # % as original Y-axis
      fill = Gender),                             # fill of bars by gender
    colour = "black",                               # black color around bars
    alpha = 1,                                      # not transparent 
    width = 0.4)+                                   # half width
  
  # flip the X and Y axes to make pyramid vertical
  coord_flip()+
  
  # manually ensure that age-axis is ordered correctly
  scale_x_discrete(limits = levels(pop_age_wpp_C_split$age))+     # defined in chunk above
  
  # set percent-axis
  scale_y_continuous(
    limits = c(min_per, max_per),                                      # min and max defined above
    breaks = seq(floor(min_per), ceiling(max_per), by = 1),                # from min% to max% by 2
    labels = paste0(                                                       # for the labels, paste together...
              abs(seq(floor(min_per), ceiling(max_per), by = 1)), "%")
    )+

  # designate colors and legend labels manually
  scale_fill_manual(
    # values = c("#7fc97f", "#beaed4")
    values = c("#1c164d", "#ec5557")
  ) +

  # plot labels, titles, caption    
  labs(
    title = "Age and gender distribution in 2020, drawn over baseline in 2005",
    x = "Age category",
    y = "Percent of total",
    fill = NULL) +
  
  # optional aesthetic themes
  theme(
    legend.position = "inside",
    legend.justification = c(1,1),
    legend.position.inside = c(1,1), 
    legend.background = element_rect(color = NA, fill = NA), 
    legend.key = element_rect(color = NA, fill=NA),
    legend.text = element_text(size=11, color = "#1c164d"),
    axis.text = element_text(size=10, color = "#1c164d"),
    axis.title = element_text(size=11, color = "#1c164d"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(color = "#1c164d", fill = NA),
    panel.background = element_rect(color = "#1c164d", fill=NA),
    axis.line = element_line(colour = "#1c164d"),
    plot.title = element_text(hjust = 0, color = "#1c164d"), 
    plot.caption = element_text(hjust=0, size=11, face = "italic"))
```

\columnbreak

\begin{center}\textcolor{fgcolor}{\textbf{\hlCeHDIRed{Key Health Indicators (latest available data)}}}\end{center}  
<!-- Maternal Mortality Ratio:   -->
<!-- Under-Five Mortality Rate   -->
\textcolor{fgcolor}{\hlCeHDIBlue{Adolescent Birth Rate:}}  
Aged 10 to 14:  `r adolescent_birth_rate_text_10_14`  
Aged 15 to 19:  `r adolescent_birth_rate_text_15_19`  

\textcolor{fgcolor}{\hlCeHDIBlue{Contraceptive Prevalence:}}  
Aged 15 to 19:  `r contraceptive_prevalence_text_15_19`  
Aged 20 to 49:  `r contraceptive_prevalence_text_20_49`  

\textcolor{fgcolor}{\hlCeHDIBlue{HPV Coverage (9-14 year old girls):}}  `r HPV_coverage_text`  
\textcolor{fgcolor}{\hlCeHDIBlue{National HPV vaccination program:}}  `r HPV_national_text`  
\textcolor{fgcolor}{\hlCeHDIBlue{UHC service coverage index:}} `r UHC_estimate |> pull(Value)` in `r UHC_estimate |> pull(YEAR)`  
\textcolor{fgcolor}{\hlCeHDIBlue{UHC Sub-Index on RMNCH:}} `r UHC_RMNCH_estimate |> pull(Value)` in `r UHC_RMNCH_estimate |> pull(YEAR)`  
\textcolor{fgcolor}{\hlCeHDIBlue{Antenatal care coverage (\ge 4 visits):}} `r antenatal_WHO_text`  
\textcolor{fgcolor}{\hlCeHDIBlue{Life expectancy at birth:}} `r life_expectancy_text`  
```{=latex}
\end{multicols}
```
<!-- \textcolor{fgcolor}{\hlCeHDIBlue{colortest}} -->
<!-- \textcolor{fgcolor}{\textbf{\hlCeHDIBlue{colortest}}} -->
<!-- \textcolor{fgcolor}{\hlCeHDIRed{colortest}} -->
<!-- \textcolor{fgcolor}{\textbf{\hlCeHDIRed{colortest}}} -->
\redrule
```{=latex}
\begin{multicols}{2}
\raggedcolumns     %new code
\interlinepenalty=10000
```
```{r, fig.width=6, fig.height=5.5}
print(params$upr_all+
        theme(
          axis.title.x = element_text(color = "black"),
          plot.background = element_rect(color = "#1c164d", fill = NA),
          panel.background = element_blank(),
          legend.background = element_blank()
        )
)
```

```{r, fig.width=4, fig.height=3.5, out.width="70%"}
print(
  params$rec_plot+
    labs(y="Recommendations (N)")+
    geom_text(aes(label = sprintf("%1.0f", n_tot), y = n_tot, vjust = -0.2), size = 5, fontface = "bold", color = "#1c164d") +
    theme(
      panel.grid = element_blank(),
      axis.text.x = element_text(size = 12, color = "#1c164d"),
      axis.text.y = element_text(size = 12, color = "#1c164d"),
      axis.title.x = element_blank(),
      axis.title.y = element_text(size = 14, color = "#1c164d"),
      legend.position = "bottom",
      # legend.position = c(0.01, 0.99),
      # legend.justification = c("left", "top"),
      legend.text = element_text(size = 12,colour = "#1c164d"),
      # plot.background = element_blank(),
      plot.background = element_rect(color = "#1c164d", fill = NA),
      panel.border = element_rect(color = "#1c164d"),
      # panel.border = element_blank(),
      axis.ticks = element_line(color = "#1c164d"),
      panel.background = element_blank(),
      legend.background = element_blank(),
      strip.background = element_blank(),
      strip.text = element_blank(),
      plot.title = element_blank()
    )
)
```
\columnbreak

```{r, fig.width=6}
print(
  params$mmr_map_plot
  # params$mmr_map_plot+theme(
  #   plot.background = element_blank()
  # )&
  # plot_annotation(theme = theme(
  #   plot.background = element_rect(color = "#1c164d", fill=NA)
  #   )
  #   )
)
```


<!-- ```{r, fig.height=3, out.width="80%"} -->
<!-- penguins |> ggplot(aes(x = island, y = bill_len, fill=species))+geom_violin() -->
<!-- ``` -->

<!-- ```{r} -->
<!-- penguins |> ggplot(aes(x = island, y = bill_len, fill=species))+geom_boxplot() -->
<!-- ``` -->

<!-- ```{r, fig.height=3, out.width="80%"} -->
<!-- penguins |> ggplot(aes(x = island, y = bill_len, fill=species))+geom_violin() -->
<!-- ``` -->

<!-- ```{r, fig.width=3, out.height="500pt", } -->
<!-- penguins |> ggplot(aes(x = island, y = bill_len, fill=species))+geom_boxplot() -->
<!-- ``` -->

<!-- ```{r, fig.height=3, out.width="80%"} -->
<!-- penguins |> ggplot(aes(x = island, y = bill_len, fill=species))+geom_violin() -->
<!-- ``` -->

<!-- ```{r} -->
<!-- penguins |> ggplot(aes(x = island, y = bill_len, fill=species))+geom_boxplot() -->
<!-- ``` -->

<!-- ```{r, fig.height=3, out.width="80%"} -->
<!-- penguins |> ggplot(aes(x = island, y = bill_len, fill=species))+geom_violin() -->
<!-- ``` -->

<!-- ```{r} -->
<!-- penguins |> ggplot(aes(x = island, y = bill_len, fill=species))+geom_boxplot() -->
<!-- ``` -->

```{=latex}
\end{multicols}
```