---
title: "Maternal Health and the UPR"
params:
  country_name: "Switzerland"
  upr_all: NULL
  rec_plot: NULL
  mmr_map_plot: NULL
  bbox_SUR_region_dynamic: NULL
  bbox_selected_SUR: NULL
  sur_area: NULL
documentclass: article
# classoption: twocolumn
output: 
  pdf_document:
    includes:
      in_header: preamble.tex
    keep_tex: true
    latex_engine: xelatex
geometry: 
  - top=2cm
  - bottom=0.5cm
  - left=0.2cm
  - right=0.2cm
papersize: a4
date: "2025-09-30"
---
```{r setup, include=FALSE}
# This chunk loads the necessary packages for the report
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.align='center', out.width="100%")
pacman::p_load(
  here,
  dplyr, forcats, ggplot2, magrittr, readr, readxl, stringr, tibble, tidyr, lubridate,
  rlang,
  ggtext, 
  patchwork,
  # apyramid,
  sf
)

sdg_data <- readRDS(here("data", "SDG_data_enhanced.rds")) |> droplevels() |> 
  mutate(response_upr = fct_recode(response_upr, "Noted" = "Noted/Other"))
state_geo <- readRDS(here("output", "state_geo_enhanced.rds"))
nearest_neighbors_list <- readRDS(here("output", "nearest_neighbors_list.rds"))
iso3_SUR <- state_geo |> filter(country==params$country_name) |> pull(iso3)

# Loop through API-generated files
for (file_name in list.files(path = here("data", "API_data"), pattern = "\\.rds$")) {
  object_name <- gsub("\\.rds$", "", file_name)
  assign(object_name, readRDS(here("data", "API_data", file_name)))
}

# UPR plot data
health_rec_C <- sdg_data |> 
  filter(state_under_review == params$country_name) |> 
      droplevels() |>
      group_by(cycle, state_under_review) |>
      count(health_related, .drop = FALSE) |> 
      group_by(cycle) |>
      mutate(
        n_tot = sum(n),
        perc = (n / n_tot) * 100,
        perc = case_when(
          health_related == "Other" ~ "",
          .default = paste0(sprintf("%1.0f", perc), "%")
        )
      ) |> ungroup()

# Life expectancy
life_expectancy <- e01dt |> 
  left_join(state_geo |> sf::st_drop_geometry() |> select(iso3,iso2, loc_id),
            join_by(country_code==loc_id)) |> 
  filter(
    !is.na(iso3),
    year>=2005
  );rm(e01dt)

pop_wpp <- pop1dt |> 
  left_join(state_geo |> sf::st_drop_geometry() |> select(iso3,iso2, loc_id),
            join_by(country_code==loc_id)) |> 
  filter(
    !is.na(iso3),
    year>=2005
  )

pop_age_wpp_C <- popAge5dt |> 
  left_join(state_geo |> sf::st_drop_geometry() |> select(iso3,iso2, loc_id),
            join_by(country_code==loc_id)) |> 
  filter(
    !is.na(iso3),
    year %in% c(2005,2020)
  ) |> 
  filter(iso3 == iso3_SUR) |> 
  # filter(year==max(year)) |> 
  select(name, year, age, popM, popF) |> 
  rename(Male=popM, Female=popF) |> 
  mutate(age = fct_inorder(age)) |> 
  pivot_longer(cols = Male:Female,
               names_to = "Gender",
               values_to = "pop"
  ) |> 
  mutate(Gender=fct(Gender)) |> 
  group_by(year) |> 
  mutate(prop = pop/sum(pop, na.rm = TRUE)) |> 
  ungroup()

fertility_rate_C <- tfr1dt |> 
  left_join(state_geo |> sf::st_drop_geometry() |> select(iso3,iso2, loc_id),
            join_by(country_code==loc_id)) |> 
  filter(
    !is.na(iso3),
    year>=2005
  ) |> 
  filter(iso3 == iso3_SUR) 

relabel_na <- function(x) {
  x <- as.character(x)
  x[is.na(x)] <- "No data"
  factor(x)
}

map_insetting <- function(
    p1, 
    plot_dat,
    bbox_SUR_region_dynamic, bbox_sur, sur_area,
    legend_title_text = "Update this text",
    p_title_text="Update this text", 
    p_caption_text = "Update this text",
    title_size=14, title_margin=17,
    caption_size=14
) {
  p2 <- p1 + 
    geom_sf(
      # data=plot_dat, 
      color="grey80", fill="transparent")+
    coord_sf(
      # expand = FALSE,
      xlim = c(bbox_SUR_region_dynamic[[1]],bbox_SUR_region_dynamic[[3]]),
      ylim = c(bbox_SUR_region_dynamic[[2]], bbox_SUR_region_dynamic[[4]])
      
      # Alternative mapping centered on the selected State
      # xlim = c(max(-180, bbox_selected_SUR()[[1]] - 20), min(180, bbox_selected_SUR()[[3]] + 20)),
      # ylim = c(max(-55.67295, bbox_selected_SUR()[[2]] - 20), min(83.6341, bbox_selected_SUR()[[4]] + 20))
    )+
    geom_rect(
      aes(
        xmin = bbox_sur["xmin"] - 0.7,
        xmax = bbox_sur["xmax"] + 0.7,
        ymin = bbox_sur["ymin"] - 0.7,
        ymax = bbox_sur["ymax"] + 0.7
      ),
      fill = "transparent",
      color = "red",
      linewidth = 0.5
    )+
    labs(fill = legend_title_text)+
    theme_void()
  
  p3 <- p1 + guides(fill="none")+
    theme_void()+
    ggnewscale::new_scale_fill()+ 
    geom_sf(
      # data=plot_dat, 
      aes(geometry=polygon, fill = selected_sur),
      color = "transparent")+ 
    scale_fill_manual(values=c("transparent", "white"))+
    # scale_linewidth_manual(values = c(0.2, 0.1)) +
    coord_sf(
      expand=FALSE,
      xlim = c(bbox_sur[[1]]-.11, bbox_sur[[3]]+.11),
      ylim = c(bbox_sur[[2]]-.11, bbox_sur[[4]]+.11)
    ) +
    # theme_void()+
    theme(
      plot.background = element_rect(fill="white"),
      panel.background = element_rect(fill="transparent"), 
      panel.border = element_rect(color = "red")
    )+
    geom_rect(
      aes(
        xmin = bbox_sur["xmin"]-.1,
        xmax = bbox_sur["xmax"]+.1,
        ymin = bbox_sur["ymin"]-.1,
        ymax = bbox_sur["ymax"]+.1
      ),
      fill = "transparent",
      color = "red",
      linewidth = 0.5
    )+
    guides(fill = "none") +
    labs(
      # subtitle = "(map zoom)", 
      caption = NULL)
  
  p_title <- plot_annotation(
    title = p_title_text,
    caption = p_caption_text,
    theme = theme(
      plot.background = element_rect(color = "#1c164d", fill=NA),
      plot.title = element_textbox_simple(
        size = title_size,
        margin = margin(t = title_margin, b = title_margin, r=0, l=0, unit = "pt")
      ),
      plot.caption = element_textbox_simple(
        size = caption_size, halign = 0.5,
        margin = margin(t = 10, b = 2, r=0, l=0, unit = "pt")
      )
    )
  )
  
  # Calculate Inset dimensions for lower left
  a<-ggplot_build(p2)
  b<-ggplot_build(p3)
  multiplier_a <- abs(a$layout$coord$limits$x[1]-a$layout$coord$limits$x[2])/abs(a$layout$coord$limits$y[1]-a$layout$coord$limits$y[2])
  multiplier_b <- abs(b$layout$coord$limits$x[1]-b$layout$coord$limits$x[2])/abs(b$layout$coord$limits$y[1]-b$layout$coord$limits$y[2])
  
  area_a <- abs(a$layout$coord$limits$x[1]-a$layout$coord$limits$x[2])*abs(a$layout$coord$limits$y[1]-a$layout$coord$limits$y[2])
  area_b <- abs(b$layout$coord$limits$x[1]-b$layout$coord$limits$x[2])*abs(b$layout$coord$limits$y[1]-b$layout$coord$limits$y[2])
  
  key_dim <- 0.3/(multiplier_a/multiplier_b)
  inset_dimensions <- if(multiplier_b<=3){c(0,0,key_dim,key_dim*multiplier_a/multiplier_b)}else{c(0,0,key_dim*multiplier_a/multiplier_b,0.2)}
  
  if (
    # sur_area > 11^11 | 
    area_b > 0.02*area_a
  ) {
    p2 + 
      theme_void() #+ p_title
  } else {
    p2 + inset_element(p3, inset_dimensions[1], inset_dimensions[2],inset_dimensions[3],inset_dimensions[4])+ 
      theme_void() #+ p_title
  }
}

```

```{r}
UHC_estimate <- UHC_all |>
  filter(IndicatorCode == "UHC_INDEX_REPORTED",
         country_name == params$country_name,
         YEAR == max(YEAR)
  )

UHC_RMNCH_estimate <- UHC_all |>
  filter(IndicatorCode == "UHC_SCI_RMNCH",
         country_name == params$country_name,
         YEAR == max(YEAR)
  )

antenatal_WHO_C <- antenatal_WHO |> filter(country_iso_3_code == iso3_SUR)
antenatal_WHO_text <- if(is_empty(antenatal_WHO_C |> 
                                  filter(year == max(year)) |> 
                                  pull(antenatal_WHO))){
  "Data not available"
} else{
  paste0(antenatal_WHO_C |> 
           filter(year == max(year)) |> 
           pull(antenatal_WHO), 
         "% in ", 
         antenatal_WHO_C |> 
           filter(year == max(year)) |> 
           pull(year)
  )
}


#########

life_expectancy_C <- life_expectancy |> filter(iso3 == iso3_SUR)
life_expectancy_text <- paste0(life_expectancy_C |> 
                                 filter(year==max(year)) |>  
                                 pull(e0B) |> round(1), 
                               " years in ",
                               life_expectancy_C |> 
                                 filter(year==max(year)) |> 
                                 pull(year),
                               " (",
                               life_expectancy_C |> 
                                 filter(year==min(year)) |>  
                                 pull(e0B) |> round(1),
                               " in ",
                               life_expectancy_C |> 
                                 filter(year==min(year)) |> 
                                 pull(year),
                               ")")

pop_C <- pop_wpp |> 
  filter(iso3==iso3_SUR)
pop_text <- paste0(round(pop_C |> filter(year==max(year)) |> pull(pop)/1000,1),
                   " million in ",
                   pop_C |> filter(year==max(year)) |> pull(year),
                   " (",
                   sprintf("%+.0f", (round(pop_C |> filter(year==max(year)) |> pull(pop)/pop_C |> filter(year==min(year)) |> pull(pop),2)-1)*100),
                   "% since ",
                   pop_C |> filter(year==min(year)) |> pull(year),
                   ")")

fertility_rate_text <- paste0(fertility_rate_C |> 
                                filter(year==max(year)) |>  
                                pull(tfr) |> round(1), 
                              " children per woman in ",
                              fertility_rate_C |> 
                                filter(year==max(year)) |> 
                                pull(year),
                              " (",
                              fertility_rate_C |> 
                                filter(year==min(year)) |>  
                                pull(tfr) |> round(1),
                              " in ",
                              fertility_rate_C |> 
                                filter(year==min(year)) |> 
                                pull(year),
                              ")")

constitutions_C <- constitutions |> filter(iso3 == iso3_SUR)
```

\maketitle
# \underline{`r params$country_name`}  

<!-- Some brief text here, for context, that runs the \textcolor{CeHDIRed}{entire width of the page}. Indicator results, tables, and figures can go into the columns below. -->
```{r, out.width="30%", warning=FALSE}
p1<-UHC_all |> 
  filter(country_name == params$country_name, 
         IndicatorCode == "UHC_INDEX_REPORTED") |>
  ggplot(aes(x = YEAR, y= NumericValue))+geom_line()+
  theme_void()

#####################
HPV_coverage_C <- HPV_coverage |> filter(SpatialDim==iso3_SUR)
HPV_coverage_text <- if(is_empty(HPV_coverage_C |> filter(YEAR == max(YEAR)) |> pull(NumericValue))){
  "Data not available"
} else{
  paste0(HPV_coverage_C |> filter(YEAR == max(YEAR)) |> pull(NumericValue), 
         "% in ", HPV_coverage_C |> filter(YEAR == max(YEAR)) |> pull(YEAR))
}


HPV_national_C <- HPV_national |> filter(SpatialDim==iso3_SUR)
HPV_national_text <- if(is_empty(HPV_national_C |> filter(YEAR == max(YEAR)) |> pull(Value))){
  "Data not available"
} else{
  paste0(HPV_national_C |> filter(YEAR == max(YEAR)) |> pull(Value), 
         " (evaluated in ", HPV_national_C |> filter(YEAR == max(YEAR)) |> pull(YEAR), ")")
}

HPV_national_alt_C <- HPV_national_alt |> filter(ISO_3_CODE==iso3_SUR)
HPV_national_alt_text <- if(is_empty(HPV_national_alt_C |> pull(INTRO_text))){
  "Data not available"
} else{
  paste0(HPV_national_alt_C |> pull(INTRO_text))
}

#####################
DTP3_coverage_C <- DTP3_coverage |> filter(SpatialDim==iso3_SUR)
DTP3_coverage_text <- if(is_empty(DTP3_coverage_C |> filter(YEAR == max(YEAR)) |> pull(NumericValue))){
  "Data not available"
} else{
  paste0(DTP3_coverage_C |> filter(YEAR == max(YEAR)) |> pull(NumericValue), 
         "% in ", DTP3_coverage_C |> filter(YEAR == max(YEAR)) |> pull(YEAR))
}

#####################
nursing_density_C <- nursing_density |> filter(SpatialDim==iso3_SUR)
nursing_density_text <- if(is_empty(nursing_density_C |> filter(YEAR == max(YEAR)) |> pull(NumericValue))){
  "Data not available"
} else{
  paste0(round(nursing_density_C |> filter(YEAR == max(YEAR)) |> pull(NumericValue),0), 
         " in ", nursing_density_C |> filter(YEAR == max(YEAR)) |> pull(YEAR))
}

####################
adolescent_birth_rate_C <- adolescent_birth_rate |> filter(SpatialDim==iso3_SUR)
adolescent_birth_rate_text_10_14 <- if(is_empty(adolescent_birth_rate_C |> 
                                                filter(Dim2=="AGEGROUP_YEARS10-14") |> 
                                                filter(YEAR == max(YEAR)) |> 
                                                pull(NumericValue))){
  "Data not available"
} else{
  paste0(adolescent_birth_rate_C |> 
           filter(Dim2=="AGEGROUP_YEARS10-14") |> 
           filter(YEAR == max(YEAR)) |> 
           pull(Value), 
         " per 1,000 women in ", 
         adolescent_birth_rate_C |> 
           filter(Dim2=="AGEGROUP_YEARS10-14") |> 
           filter(YEAR == max(YEAR)) |> 
           pull(YEAR)
  )
}
adolescent_birth_rate_text_15_19 <- if(is_empty(adolescent_birth_rate_C |> 
                                                filter(Dim2=="AGEGROUP_YEARS15-19") |> 
                                                filter(YEAR == max(YEAR)) |> 
                                                pull(NumericValue))){
  "Data not available"
} else{
  paste0(adolescent_birth_rate_C |> 
           filter(Dim2=="AGEGROUP_YEARS15-19") |> 
           filter(YEAR == max(YEAR)) |> 
           pull(Value), 
         " per 1,000 women in ", 
         adolescent_birth_rate_C |> 
           filter(Dim2=="AGEGROUP_YEARS15-19") |> 
           filter(YEAR == max(YEAR)) |> 
           pull(YEAR)
  )
}

##############
contraceptive_prevalence_C <- contraceptive_prevalence |> 
  filter(Dim1Type == "AGEGROUP") |> 
  filter(SpatialDim==iso3_SUR)

contraceptive_prevalence_text_15_19 <- 
  if(is_empty(contraceptive_prevalence_C |> 
              filter(Dim1=="AGEGROUP_YEARS15-19") |> 
              filter(YEAR == max(YEAR)) |> 
              pull(NumericValue))){
    "Data not available"
  } else{
    paste0(round(contraceptive_prevalence_C |> 
                   filter(Dim1=="AGEGROUP_YEARS15-19") |> 
                   filter(YEAR == max(YEAR)) |> 
                   pull(NumericValue),1), 
           "% in ", 
           contraceptive_prevalence_C |> 
             filter(Dim1=="AGEGROUP_YEARS15-19") |> 
             filter(YEAR == max(YEAR)) |> 
             pull(YEAR)
    )
  }
contraceptive_prevalence_text_20_49 <- 
  if(is_empty(contraceptive_prevalence_C |> 
              filter(Dim1=="AGEGROUP_YEARS20-49") |> 
              filter(YEAR == max(YEAR)) |> 
              pull(NumericValue))){
    "Data not available"
  } else{
    paste0(round(contraceptive_prevalence_C |> 
                   filter(Dim1=="AGEGROUP_YEARS20-49") |> 
                   filter(YEAR == max(YEAR)) |> 
                   pull(NumericValue),1),
           "% in ", 
           contraceptive_prevalence_C |> 
             filter(Dim1=="AGEGROUP_YEARS20-49") |> 
             filter(YEAR == max(YEAR)) |> 
             pull(YEAR)
    )
  }

```
```{=latex}
\begin{multicols}{2}
\raggedcolumns     %new code
\interlinepenalty=10000
```
\begin{center}\textcolor{fgcolor}{\textbf{\hlCeHDIRed{\ \ \ \ The Right to Health \ \ \ \ }}}\end{center}  
As enshrined in Article 12 of the International Convenant on Economic, Social and Cultural Rights, the right to health is an inclusive human right that extends beyond timely and appropriate health care to encompass the underlying determinants of health.  

Under the Right to Health, States have the following obligations:  
\ \ - \textcolor{fgcolor}{\hlCeHDIBlue{Respect:}} refrain from directly or indirectly interfering with the enjoyment of the right to health.  
\ \ - \textcolor{fgcolor}{\hlCeHDIBlue{Protect:}} take effective measures to prevent third parties from undermining of violating the guarantees of the right to health.  
\ \ - \textcolor{fgcolor}{\hlCeHDIBlue{Fulfill:}} adopt appropriate legislative, administrative, budgetary, judicial, promotional, and other measures toward the full realization of the right to health.
\begin{center}\textcolor{fgcolor}{\textbf{\hlCeHDIRed{\ \ \ \ The Universal Periodic Review (UPR) \ \ \ \ }}}\end{center}  
The UPR is a **State-led mechanism** that aims to evaluate each State’s human rights obligations and commitments. The review process is participatory and includes interactive discussions during which any UN Member State may issue recommendations to the State under review, which may then choose to **"Support"** or **"Note"** those recommendations.  

A CeHDI analysis of recommendations from all cycles of the UPR was used to identify health-related recommendations, which have globally gained prominence in the UPR over time (for `r params$country_name`, health-related recommendations accounted for `r health_rec_C |> filter(cycle == "Cycle 1", health_related == "Health-related") |> pull(perc)` of all recommendations in Cycle 1, and `r health_rec_C |> filter(health_related == "Health-related") |> slice_max(order_by = cycle) |> pull(perc)` in `r health_rec_C |> filter(health_related == "Health-related") |> slice_max(order_by = cycle) |> pull(cycle)`). These recommendations were further categorized into the following themes:\footnote{Visit \href{https://cehdi-har.share.connect.posit.cloud/}{cehdi-har.share.connect.posit.cloud} for more details.\label{fn:cehdiHAR}}
```{r, fig.width=5.2, fig.height=4}
print(params$upr_all+
        labs(title = params$country_name)+
        theme(
          axis.title.x = element_text(color = "#1c164d"),
          axis.text = element_text(color = "#1c164d"),
          plot.title = element_text(color = "#1c164d"),
          plot.caption = element_text(color = "#1c164d"),
          plot.background = element_rect(color = "#1c164d", fill = NA),
          panel.background = element_blank(),
          legend.background = element_blank()
        )
)
```
\columnbreak
\begin{center}\textcolor{fgcolor}{\textbf{\hlCeHDIRed{\ \ \ \ Demographics \ \ \ \ }}}\end{center}  
\textcolor{fgcolor}{\hlCeHDIBlue{Population:}}\footnote{\href{https://population.un.org/wpp}{population.un.org/wpp} \label{fn:WPP}} `r pop_text`  
\textcolor{fgcolor}{\hlCeHDIBlue{Total fertility rate:}}\textsuperscript{\ref{fn:WPP}} `r fertility_rate_text`  

\begin{center}\textcolor{fgcolor}{\textbf{\hlCeHDIRed{\ \ \ \ Key Health Indicators (latest available data) \ \ \ \ }}}\end{center}  
\textcolor{fgcolor}{\hlCeHDIBlue{Maternal mortality ratio (MMR):}}\textsuperscript{\ref{fn:whoGHO}} `r MMR |> filter(SpatialDim == iso3_SUR, YEAR == 2023) |> pull(NumericValue) |> round(digits = 0)` maternal deaths per 100,000 live births in 2023 (`r MMR |> filter(SpatialDim == iso3_SUR, YEAR == 2005) |> pull(NumericValue) |> round(digits = 0)` in 2005)  
```{r, }
# mmr_estimate_2023 = params$mmr_map_plot@data |>
#   filter(country_name == params$country_name,
#          YEAR == "2023") |>
#   pull(NumericValue) |>
#   round(0)
# 
# map_insetting(
#   p1 = params$mmr_map_plot,
#   p_caption_text = paste0(params$country_name, ": ", mmr_estimate_2023, " per 100,000 live births"),
#   legend_title_text = "MMR",
#   p_title_text = "Maternal mortality ratio (MMR) estimates in 2023",
#   bbox_SUR_region_dynamic = params$bbox_SUR_region_dynamic,
#   bbox_sur = params$bbox_selected_SUR,
#   sur_area = params$sur_area
# )+labs(fill="MMR")

p1<-MMR |> 
  filter(YEAR==2023, !is.na(country_name)) |> 
  right_join(state_geo, by=join_by(SpatialDim==iso3)) |> 
  mutate(selected_sur = factor(case_when(
    country_name == params$country_name ~ params$country_name,
    .default = "Other"
  ),
  levels = c(params$country_name, "Other")
  )) |> 
  ggplot(aes(geometry = polygon))+
  geom_sf(aes(fill=mmr_cat), color = NA)+
  scale_fill_brewer(palette = "YlOrRd", na.value = "grey80", labels = relabel_na)+
  theme_void()+
  labs(fill="MMR\nin 2023")

p2 <- p1 +
  geom_sf(color="grey70", fill=NA)+
  geom_rect(
    aes(
      xmin = params$bbox_selected_SUR["xmin"] - 0.7,
      xmax = params$bbox_selected_SUR["xmax"] + 0.7,
      ymin = params$bbox_selected_SUR["ymin"] - 0.7,
      ymax = params$bbox_selected_SUR["ymax"] + 0.7
    ),
    fill = "transparent",
    color = "red",
    linewidth = 0.5
  )+
  coord_sf(
    # expand = FALSE,
    xlim = c(params$bbox_SUR_region_dynamic[[1]],params$bbox_SUR_region_dynamic[[3]]),
    ylim = c(params$bbox_SUR_region_dynamic[[2]], params$bbox_SUR_region_dynamic[[4]])
  )

p3<-p1+
  guides(fill="none")+
  ggnewscale::new_scale_fill()+ 
  geom_sf(
    aes(fill = selected_sur),
    color = "transparent")+ 
  scale_fill_manual(values=c("transparent", "white"))+
  guides(fill="none")+
  coord_sf(
    expand=FALSE,
    xlim = c(params$bbox_selected_SUR[[1]]-.11, params$bbox_selected_SUR[[3]]+.11),
    ylim = c(params$bbox_selected_SUR[[2]]-.11, params$bbox_selected_SUR[[4]]+.11)
  )+
  theme(panel.background = element_rect(fill="white"))+
  geom_rect(
    aes(
      xmin = params$bbox_selected_SUR["xmin"]-.1,
      xmax = params$bbox_selected_SUR["xmax"]+.1,
      ymin = params$bbox_selected_SUR["ymin"]-.1,
      ymax = params$bbox_selected_SUR["ymax"]+.1
    ),
    fill = "transparent",
    color = "red",
    linewidth = 0.5
  )

a<-ggplot_build(p2)
b<-ggplot_build(p3)
multiplier_a <- abs(a$layout$coord$limits$x[1]-a$layout$coord$limits$x[2])/abs(a$layout$coord$limits$y[1]-a$layout$coord$limits$y[2])
multiplier_b <- abs(b$layout$coord$limits$x[1]-b$layout$coord$limits$x[2])/abs(b$layout$coord$limits$y[1]-b$layout$coord$limits$y[2])

area_a <- abs(a$layout$coord$limits$x[1]-a$layout$coord$limits$x[2])*abs(a$layout$coord$limits$y[1]-a$layout$coord$limits$y[2])
area_b <- abs(b$layout$coord$limits$x[1]-b$layout$coord$limits$x[2])*abs(b$layout$coord$limits$y[1]-b$layout$coord$limits$y[2])

key_dim <- 0.3/(multiplier_a/multiplier_b)
inset_dimensions <- if(multiplier_b<=3){c(0,0,key_dim,key_dim*multiplier_a/multiplier_b)}else{c(0,0,key_dim*multiplier_a/multiplier_b,0.2)}

if (
  # sur_area > 11^11 | 
  area_b > 0.02*area_a
) {
  p2
} else {
  p2 + inset_element(p3, inset_dimensions[1], inset_dimensions[2],inset_dimensions[3],inset_dimensions[4])+ plot_annotation(
    theme = theme(plot.margin = margin(0,0,0,0), plot.background = element_blank()
                  ))
}

# params$mmr_map_plot

# params$mmr_map_plot+theme(
#   plot.background = element_blank()
# )&
# plot_annotation(theme = theme(
#   plot.background = element_rect(color = "#1c164d", fill=NA)
#   )
#   )
```
\textcolor{fgcolor}{\hlCeHDIBlue{Life expectancy at birth:}}\textsuperscript{\ref{fn:WPP}} `r life_expectancy_text`
<!-- Maternal Mortality Ratio:   -->
<!-- Under-Five Mortality Rate   -->  
\textcolor{fgcolor}{\hlCeHDIBlue{Adolescent birth rate:}}\footnote{\href{https://www.who.int/data/gho/info/gho-odata-api}{WHO Global Health Observatory (accessed through API)} \label{fn:whoGHO}}  
\ \ - Aged 10 to 14: `r adolescent_birth_rate_text_10_14`  
\ \ - Aged 15 to 19: `r adolescent_birth_rate_text_15_19`  

\textcolor{fgcolor}{\hlCeHDIBlue{Contraceptive prevalence (modern methods):}}\textsuperscript{\ref{fn:whoGHO}}  
\ \ - Aged 15 to 19:  `r contraceptive_prevalence_text_15_19`  
\ \ - Aged 20 to 49:  `r contraceptive_prevalence_text_20_49`
\begin{center}\textcolor{fgcolor}{\textbf{\hlCeHDIRed{\ \ \ \ Health Services Coverage (latest available data) \ \ \ \ }}}\end{center}  
\textcolor{fgcolor}{\hlCeHDIBlue{UHC service coverage index:}}\textsuperscript{\ref{fn:whoGHO}} `r UHC_estimate |> pull(Value)` in `r UHC_estimate |> pull(YEAR)`  
\textcolor{fgcolor}{\hlCeHDIBlue{UHC sub-Index on RMNCH:}}\textsuperscript{\ref{fn:whoGHO}} `r UHC_RMNCH_estimate |> pull(Value)` in `r UHC_RMNCH_estimate |> pull(YEAR)`  
\textcolor{fgcolor}{\hlCeHDIBlue{Antenatal care coverage (\ge 4 visits):}}\footnote{\href{https://platform.who.int/data/maternal-newborn-child-adolescent-ageing}{WHO MNCAH Data Portal} \label{fn:whoMNCAH}} `r antenatal_WHO_text`  
\textcolor{fgcolor}{\hlCeHDIBlue{Nursing/midwifery personnel (per 10,000):}}\textsuperscript{\ref{fn:whoGHO}}  `r nursing_density_text`  

\textcolor{fgcolor}{\hlCeHDIBlue{HPV coverage (9-14 year old girls):}}\textsuperscript{\ref{fn:whoGHO}}  `r HPV_coverage_text`  
\textcolor{fgcolor}{\hlCeHDIBlue{HPV vaccine introduction:}}\footnote{\href{https://immunizationdata.who.int}{immunizationdata.who.int} \label{fn:whoimmunization}} `r HPV_national_alt_text`  
\textcolor{fgcolor}{\hlCeHDIBlue{DTP3 coverage (among 1-year-olds):}}\textsuperscript{\ref{fn:whoGHO}} `r DTP3_coverage_text`  
\begin{center}\textcolor{fgcolor}{\textbf{\hlCeHDIRed{\ \ \ \ Legal Basis for Health \ \ \ \ }}}\end{center}  
\textcolor{fgcolor}{\hlCeHDIBlue{Constitutional guarantees for citizens (as of June 2024):}}\footnote{\href{https://www.worldpolicycenter.org/maps-data/data-download/constitutions-data-download}{worldpolicycenter.org} \label{fn:worldpolicy}}  
\ \ - Right to health: `r constitutions |> filter(iso3 == iso3_SUR) |> pull(const_health)`  
\ \ - Right to public health: `r constitutions |> filter(iso3 == iso3_SUR) |> pull(const_publichealth)`  
\ \ - Right to medical care: `r constitutions |> filter(iso3 == iso3_SUR) |> pull(const_medcare)`  
\ \ - Right to a healthy environment: `r constitutions |> filter(iso3 == iso3_SUR) |> pull(const_enviro)`  
\textcolor{fgcolor}{\hlCeHDIBlue{Legal status of abortion (as of June 2023):}}\footnote{\href{https://reproductiverights.org/wp-content/uploads/2023/12/WALM_2023-v3-Updated_12-20-23.pdf}{reproductiverights.org} \label{fn:reproductiverights}} `r world_abortion_laws |> filter(country == params$country_name) |> pull(category)`  
```{=latex}
\end{multicols}
```