---
title: "Definitions of recommendations related to the Right to Health in the Universal Periodic Review"
subtitle: "Recommendations downloaded from UHRI"
author: "Anshu Uppal"
date: "2025-10-06"
date-modified: last-modified # Dynamic modified date
format:
        html:
                code-fold: true
                toc: true
                toc-location: left  # Place TOC on the left side
                toc-depth: 5        # Include headers up to level 3 (###)
                toc-title: "On this page" # Custom title for the TOC
                smooth-scroll: true # Enable smooth scrolling
                embed-resources: true # Make the html file self contained
                grid: 
                  body-width: 1100px
                  margin-width: 100px
reference-location: margin
citation-location: margin
fig-format: svg
knitr: 
  opts_chunk: 
    dev: svglite
    fig.path: output/figures/
---

## Setup
```{r}
# install.packages("pacman")
pacman::p_load(
  here,
  tidyverse,
  # plotly,
  janitor,
  DT,
  sf,
  necountries,
  patchwork,
  Hmisc,
  ggtext
)

# Load or install packages from GitHub:
# pacman::p_load_gh(
#   
# )

# Load in custom functions
# source(here("utils.R"))
```

### Download and pre-process the UPR data from UHRI  
Download the full UHRI dataset from <https://uhri.ohchr.org/en/our-data-api>
The direct download of the excel file is: <https://dataex.ohchr.org/uhri/export-results/export-full-en.xlsx>
```{r}
# state_geo <- readRDS(here("output", "state_geo_enhanced.rds"))
# 
# httr::GET("https://dataex.ohchr.org/uhri/export-results/export-full-en.xlsx", httr::write_disk(tf <- tempfile(fileext = ".xlsx")))
# df_0 <-  read_xlsx(tf,
#                    # Specify the column types, otherwise some show up as NA in R
#                    col_types = c(
#                      rep("text", 7), "date", rep("text", 7), "date")) |>
#   janitor::clean_names();unlink(tf);rm(tf)
# df<-df_0 |> janitor::clean_names() |>
#   # Add the date that the dataset was accessed from the UHRI website
#   mutate(date_accessed = ymd(Sys.Date())) |>
#   mutate(countries_concerned = str_remove(countries_concerned, "- "),
#          reccomending_body = str_remove(reccomending_body, "- ")
#   ) |>
#   mutate(
#     text1 = text,
#     # text = str_remove_all(text, "&nbsp;"),
#     # text = str_remove_all(text, "&nbsp"),
#     text = str_replace_all(text, "’", "'"),
#     text = str_replace_all(text, "“", "'"),
#     text = str_replace_all(text, "”", "'"),
#     text = str_replace_all(text, "\\( ", "\\("),
#     text = str_replace_all(text, " \\)", "\\)"),
#     text = str_replace_all(text, " ' s ", "'s"),
#     text = str_replace_all(text, "' s ", "'s"),
#     text = str_replace_all(text, "&nbsp;", " "),
#     text = str_replace_all(text, "&nbsp", " "),
#     text = str_replace_all(text, "&rsquo;", "'"),
#     text = str_replace_all(text, "&ndash;|&#8209;", "-"),
#     text = str_replace_all(text, "&ocirc;", "ô"),
#     text = str_replace_all(text, "&ldquo;|&rdquo;", "'"),
#     text = str_replace_all(text, "&uacute;", "ú"),
#     text = str_replace_all(text, "&uuml;", "ü"),
#     text = str_replace_all(text, "&amp;", "&"),
#     text = str_replace_all(text, "&deg;", "°"),
#     text = str_replace_all(text, "&eacute;", "é"),
#     text = str_replace_all(text, "&egrave;", "è"),
#     text = str_replace_all(text, "&lrm;", ""),
#     text = str_replace_all(text, "&iacute;", "í"),
#     
#     text = str_squish(text),
#     text = case_when(str_starts(text, "R ") ~ str_remove(text, "R "),
#                      .default = text),
#     # text = str_replace(text, " . ", ". "),
#     # text = str_replace(text, ":.", ":"),
#     # text = str_replace(text, ",.", ":"),
#     text = str_remove_all(text, "\\[...\\]")
#   ) |>
#   arrange(countries_concerned, reccomending_body, document_publication_date) |>
#   rename(document = document_symbol, body = reccomending_body) |>
#   mutate(
#     paragraph = case_when(
#       # CASE 1: The text starts with "Recommendation No."
#       # This detects the first pattern you described.
#       # str_detect(text, "^Recommendation No\\.") ~ {
#       #   # str_match() lets us capture parts of the text.
#       #   # We look for "para. [number]: [number]"
#       #   matches <- str_match(text, "para\\.\\s*(\\d+):\\s*(\\d+)")
#       #   # We then take the captured numbers (the 2nd and 3rd elements from the match)
#       #   # and paste them together with a dot, creating the "96.21" format.
#       #   paste(matches[, 2], matches[, 3], sep = ".")
#       # },
#       str_detect(text, "^Recommendation No") ~ {
#         # This regex now finds the paragraph number, skips to the recommendation number, and captures both.
#         # It works whether there is a colon or not.
#         matches <- str_match(text, "para\\.\\s*(\\d+).*?(\\d+)\\.")
#         paste(matches[, 2], matches[, 3], sep = ".")
#       },
#       
#       # NEW CASE: The text starts with "Recommendations contained in para."
#       # str_detect(text, "^Recommendations contained in para") ~ str_match(text, "para\\.\\s*(\\d+)")[, 2],
#       # str_detect(text, "^Recommendation(s)? contained in para(graph)?") ~ str_match(text, "[Pp]ara(?:graph)?\\s*(\\d+\\(?[a-z]?\\)?)")[, 2],
#       str_detect(text, "^Recommendation(s)? (?:contained in|from) para(graph)?") ~ str_match(text, "[Pp]ara(?:graph)?\\s*(.*?):")[, 2],
#       str_detect(text, "^Recommendation(s)? contained in ") ~ str_match(text, "contained in\\s*(\\d+\\(?[a-z]?\\)?)")[, 2],
#       
#       # CASE 2: The text starts with "para" or "paragraph"
#       # This is the new case to handle paragraph-only identifiers.
#       str_detect(text, "^[Pp]ara") ~ str_match(text, "[Pp]ara(?:graph)?\\.?\\s*(\\d+)")[, 2],
#       
#       # CASE 3: The text starts with a "number.number" format (e.g., "136.1 ...")
#       # This must come before the single-number check to be matched correctly.
#       str_detect(text, "^\\d+\\.\\d+") ~ str_extract(text, "^\\d+\\.\\d+"),
#       
#       # CASE 4: The text starts with just one number (e.g., "1 ...")
#       # This is the new case to handle single leading numbers.
#       str_detect(text, "^\\d+") ~ str_extract(text, "^\\d+"),
#       
#       # If none of the above patterns match, return NA
#       TRUE ~ NA_character_
#     ),
#     upr_session = str_remove(upr_session, "- "),
#     upr_session = str_split_i(upr_session, " - ", 1),
#     upr_session_number = as.numeric(str_extract(upr_session, "\\d+")),
#     upr_cycle = case_when(
#       upr_session_number <= 12 ~ "Cycle 1",
#       upr_session_number <= 26 ~ "Cycle 2",
#       upr_session_number <= 40 ~ "Cycle 3",
#       upr_session_number <= 54 ~ "Cycle 4",
#     ),
#     type = str_remove(type, "- "),
#     upr_position = str_remove(upr_position, "- "),
#     title_a = str_split_i(paragraph,"\\.",1),
#     title_a = case_when(is.na(as.numeric(title_a)) ~ str_extract(title_a, "^\\d+"),
#                         .default = title_a),
#     title_b = str_pad(str_split_i(paragraph,"\\.",2), , width = 3, side = "left", pad = "0"),
#     title_b = case_when(is.na(title_b) | str_starts(title_b, "\\(") ~ "0",
#                         .default = title_b),
#     title_2 = as.numeric(paste(title_a, title_b,sep = "."))
#   )|>
#   arrange(upr_session_number) |>
#   mutate(
#     upr_session = fct_inorder(upr_session),
#     upr_cycle = fct_inorder(upr_cycle)
#   ) |>
#   # select(-title_a, -title_b) |>
#   relocate(paragraph, .after=document) |>
#   relocate(title_2, upr_position, type, .after = paragraph) |>
#   arrange(countries_concerned, body, document_publication_date, title_2, paragraph) |>
#   rename(
#     state_under_review = countries_concerned,
#     document_code = document,
#     session = upr_session,
#     cycle = upr_cycle,
#     recommending_state_upr = upr_reccomending_states,
#     response_upr = upr_position
#   ) |>
#   left_join(state_geo |> select(country, iso3, iso2) |> st_drop_geometry(),
#             by = join_by(state_under_review == country))
# 
# 
# # Filter for only the UPR dataset
# upr_df <- df |> filter(body == "UPR") |> droplevels() |>
#   mutate(
#     recommending_state_upr = str_remove(recommending_state_upr, "^- "),
#     recommending_states_count = case_when(
#       is.na(recommending_state_upr) ~ NA,
#       .default = str_count(recommending_state_upr, "-") %>% coalesce(0L)+1
#     ),
#     text_2 = case_when(
#       # CASE 1: Remove "Recommendation No. ..." prefix
#       str_detect(text, "^Recommendation No") ~ str_remove(text, "^Recommendation No.*?\\d+\\.\\s*"),
#       # str_detect(text, "^Recommendations contained in para") ~ str_remove(text, "^Recommendations contained in para\\.\\s*\\d+:\\s*"),
#       str_detect(text, "^Recommendation(s)? contained in para(graph)?") ~ str_remove(text, "^Recommendation(?:s)? contained in para(?:graph)?.*?\\:\\s*"),
#       str_detect(text, "^Recommendation(s)? contained in ") ~ str_remove(text, "^Recommendation(?:s)? contained in .*?\\:\\s*"),
#       
#       # CASE 2: Remove "para. ..." prefix
#       str_detect(text, "^[Pp]ara") ~ str_remove(text, "^[Pp]ara(?:graph)?\\.?\\s*(\\d+\\.?)\\s*"),
#       # CASE 3: Remove "number.number" prefix
#       str_detect(text, "^\\d+\\.\\d+") ~ str_remove(text, "^\\d+\\.\\d+\\.?\\s*"),
#       # CASE 4: Remove single number prefix
#       str_detect(text, "^\\d+") ~ str_remove(text, "^\\d+\\.?\\s*"),
#       # If no pattern matches, just keep the original text
#       TRUE ~ text
#     ),
#     text_2 = case_when(
#       str_detect(text_2, "^\\d+") ~ str_remove(text_2, "^\\d+\\.?\\s*"),
#       TRUE ~ text_2
#     ),
#     text_2 = case_when(
#       str_detect(text_2, "^\\d+") ~ str_remove(text_2, "^\\d+\\.?\\s*"),
#       TRUE ~ text_2
#     ),
#     text_2 = str_remove(text_2,"^\\."),
#     text_2 = str_remove(text_2,"^\\:"),
#     text_2 = str_remove(text_2,"^\\;"),
#     text_2 = str_squish(text_2),
#     # text_2 = str_replace(text_2, ". visit", "a visit"),
#     # text_2 = str_replace(text_2, ". standing", "a standing"),
#     # text_2 = str_replace(text_2, ". method", "a method"),
#     # text_2 = str_replace(text_2, ". dialogue", "a dialogue"),
#     # text_2 = str_replace(text_2, ". comprehensive", "a comprehensive"),
#     # text_2 = str_replace(text_2, ". national", "a national")
#   ) |>
#   filter(!str_starts(text, "Response of "),
#          response_upr != "Concerns/Observations") |>
#   rowid_to_column() |>
#   
#   # Create a new column that lists recommending states by commas
#   mutate(recommending_state_upr_comma = recommending_state_upr) |>
#   separate_wider_delim(recommending_state_upr_comma, delim = "-", names_sep = "_", too_few = "align_start") |>
#   mutate(across(starts_with("recommending_state_upr_comma"), ~ str_squish(.))) |>
#   unite(col = "recommending_state_upr_comma", sep = ", ", starts_with("recommending_state_upr_comma"), na.rm = TRUE) |>
#   relocate(text1, text_2, .after = text) |>
#   relocate(recommending_state_upr_comma, recommending_states_count, .after= recommending_state_upr) |>
#   mutate(recommending_state_upr_comma = case_when(recommending_state_upr_comma == "" ~ NA, .default = recommending_state_upr_comma)) |>
#   mutate(response_upr = factor(
#     case_when(
#       response_upr == "Supported/Noted" ~ "Partially supported",
#       .default = response_upr),
#     levels = c("Supported", "Partially supported", "Noted")
#   )
#   )
# 
# saveRDS(df, here("data", "UHRI_full.rds"))
# saveRDS(df, here("data", paste0("UHRI_full_", Sys.Date(), ".rds")))
# saveRDS(upr_df, here("data", "UHRI_UPR.rds"))
# saveRDS(upr_df, here("data", paste0("UHRI_UPR_", Sys.Date(), ".rds")))
```


```{r}
# Read in the UPR data
uhri_upr <- readRDS(here("data", "UHRI_UPR.rds"))
```


For the present analysis, the dataset was downloaded on `r unique(uhri_upr$date_accessed)`.  

## Create definitions 

**Still to refine, especially the definitions related to maternal health**

See also: <https://www.who.int/health-topics>

### Right to health
I then created a set of keywords to assign recommendations as being related to the "right to health". This is a key step to refine in case we opt for an automated process. Some of the keywords are partial words because I use a partial word matching (e.g. "prophyla" will match with both "prophylaxis" as well as "prophylactics"). **Need to be careful to not accidentally include irrelevent terms.**

Need to review inclusion of remaining recommendations with reference to CEDAW (~400 entries)
```{r}
# Define keywords to use to identify recommendations related to the right to health
health_keywords <- "health|sanitation|sanitary|safe water|clean water|right to water|disease|sick|infectio|nutrition|tuberculosis|malaria|with hiv| hiv |hiv/aids|hiv-aids|hiv and aids|hiv-positive| sti | tb |hospital|clinic|vaccin|immunisation|immunization|virus|viral|medic|doctor|nurse|nursing|maternal|contracep|abortion|reproductive|pregnan|prenatal|postnatal|neonatal|breastfe|obstetric|fertility|medic|diabetes|cancer|blood|hypertension|respiratory|prophyla|violence against women|violence against children|domestic violence|family violence|violence in the family|gender-based violence|sexual violence|marital violence|gender violence| rape|virginity|hunger|malnutrition|obesity|sexual abuse|mortality|leprosy|famine|drought|food|tobacco|smoking|child marriage|forced marriage|early marriage|infanticide|mutilation|honor crime|honor killing|honour crime|honour killing|breast ironing|harmful practices|harmful traditional practices|epidem|sterilization|contracept|sex education|suicide|cedaw|femicide|polygam|covid|pandemic|corona|palliative"

# "well-being|wellbeing|disaster|disabled|disabili"

NOT_health_keywords <- "blood feud"

# Define combinations of keywords for the right to health
keywords_comb1a <- "child|girl|women|sexual|domestic|gender|marital|lgbt|same-sex"
keywords_comb1b <- "abuse|maltreatment|violence|sexual"

keywords_comb2a <- "forced|minimum"
keywords_comb2b <- "marriage"

keywords_comb3a <- "sex|gender|civil identity|transgender"
keywords_comb3b <- "surgery|sterili"


```

#### Health systems and services
> In 2007 the World Health Organization (WHO) proposed a framework describing health systems in terms of six core components or ‘building blocks’: (i) service delivery and safety; (ii) health workforce; (iii) health statistics and information systems; (iv) access to essential medicines; (v) financing; and (vi) leadership/ governance.  

When applying these definitions, I think we will need to first more generally define health-related recommendations, and then within that context we can apply the keywords below to define health systems and services.

```{r}
hss_keywords <- "health worker|healthcare worker|health-care worker|uhc|universal health|health system|healthcare system|health-care system|health care system|health service|healthcare service|health-care service|health care service| oral health|health and education|health coverage|public health|free health|access to health|quality health|strenghthen health|right to health"
hss_other_keywords <- "data"

hss_keywords_comb1a <- "health"
hss_keywords_comb1b <- "capacity-building|coverage|infrastructure|human resources|universal access|affordab|facilities| access | public |budget"

hss_keywords_comb2a <- "health"
hss_keywords_comb2b <- "sector|system"
hss_keywords_comb2c <- "financ"

```

#### Health security, emergencies, and disaster relief
> This category relates to health in the context of humanitarian, emergency and disaster relief in order to reduce avoidable loss of life and the burden of disease and disability. Emergencies include any event that may have negative consequences for human health including events that have not yet led to disease in humans but have the potential to cause human disease through exposure to infected or contaminated food, water, animals, manufactured products or environments. Emergencies are situations impacting the lives and well-being of a large number of people or significant percentage of a population and requiring substantial multi-sectoral assistance. For WHO to respond, there must be clear health consequences.

```{r}
emergencies_keywords <- "covid|pandemic|epidemic|drought"

emergencies_keywords_comb1a <- "disaster"
emergencies_keywords_comb1b <- "health|life"

emergencies_keywords_comb2a <- "earthquake|flood|drought"
emergencies_keywords_comb2b <- "relief"
```

#### NCDs
> The four main types of noncommunicable diseases (NCD) are cardiovascular diseases (like heart attacks and stroke); cancers; chronic respiratory diseases (such as chronic obstructed pulmonary disease and asthma); diabetes; alcohol and drug use. But this category also includes other types of NCD.

```{r}
ncd_keywords <- "non-communicable|cardiovascular|stroke|cancer|neoplasm|chronic diseas|hypertension|diabet|alcohol|tobacco|smoking|kidney| liver |air pollution|drug dependence"

ncd_keywords_comb1a <- " drug"
ncd_keywords_comb1b <- "addict|abuse| use |harm"

ncd_keywords_comb2a <- " air "
ncd_keywords_comb2b <- "pollution| quality"

ncd_exclusion <- ""
```
#### Communicable diseases
> Communicable diseases comprise both infectious diseases and zoonotic diseases. Infectious diseases are caused by pathogenic microorganisms, such as bacteria, viruses, parasites or fungi; the diseases can be spread, directly or indirectly, from one person to another. Zoonotic diseases are infectious diseases of animals that can cause disease when transmitted to humans.

```{r}
communicable_keywords <- " communicable|infectious|zoonotic|pathogen|parasit|fungal|ebola|measles|influenza|polio|norovirus|vector|water-borne|trachoma"

communicable_keywords_comb1a <- "virus"
# In this instance, exempt combinations with the following (these will fall purely under HIV/AIDs):
communicable_keywords_comb1b <- "with hiv| hiv |hiv/aids|hiv-aids|hiv and aids|hiv-positive| hiv,| hiv-"

communicable_exclusion <- ""
```

#### SRHR
> Sexual health is a state of physical, mental and social well-being in relation to sexuality. It requires a positive and respectful approach to sexuality and sexual relationships, as well as the possibility of having pleasurable and safe sexual experiences, free of coercion, discrimination and violence.  

> Reproductive health is concerned with a state of physical, mental and social well-being in relation to sexuality, as well as the reproductive processes, functions and system at all stages of life. Examples include the freedom to decide if, when, and how often to engage in sexual activity; the right of men and women to be informed about and have access to safe, effective, affordable and acceptable methods of fertility regulation of their choice; and the right of women to access appropriate health care services that will enable them to go through pregnancy safely.

May need to first define "Health of LGBTI persons", and then add logic so that other issues related to LGBTI rights, and not included in the health of LGBTI classification, will fall under SRHR.
```{r}
srhr_keywords <- "reproductive|sex education|sexuality education|education on sexual|breastfe|fertility|sterilization|sexual health| sexual education"

srhr_keywords_comb1a <- "sex|gender|civil identity|transgender"
srhr_keywords_comb1b <- "sterili"

srhr_keywords_comb2a <- "sexual relation|sexual conduct|same-sex|gay|lesbian|same sex|homosexual"
srhr_keywords_comb2b <- "criminal"#|prohibit"
```

#### Mental health
> Mental health refers to a broad array of activities directly or indirectly related to the mental well-being component included in the WHO’s definition of health: “A state of complete physical, mental and social well-being, and not merely the absence of disease”. It is related to the promotion of well-being, the prevention of mental disorders, and the treatment and rehabilitation of people affected by mental disorders.

```{r}
mental_keywords <- "anxiety|depression| mental|suicide|psychosocial|psychiatr|involuntary hospital|forced hospital"

mental_keywords_comb1a <- " mental"
mental_keywords_comb1b <- "health|disorder|illness"
```
#### Social and economic determinants of health
> The social determinants of health (SDH) are the conditions in which people are born, grow, work, live, and age, and the wider set of forces and systems shaping the conditions of daily life. These forces and systems include economic policies and systems, development agendas, social norms, social policies and political systems. The Sustainable Development Goals (SDGs) provide a comprehensive blueprint for human development and for systematically addressing the social determinants of health.

May need to first generally define health, and then classify SOCED as health-related recommendations that touch on social determinants
```{r}
SOCED_keywords <- "social determinants|migrant|poverty| poor|destitut|inequality|vulnerable| elder| older"
```

#### GBV - Gender-based violence
> This category includes any violence which has been perpetrated against another based on their sex, gender identity and expression, or their sexual orientation, as well as harmful practices such as female genital mutilation (FGM), female infanticide, sex-selective abortions, forced and early marriage, honour crimes or killings, polygamy, wife inheritance, virginity tests, breast ironing and similar practices.

```{r}
GBV_keywords <- "child marriage|early marriage|forced marriage|marriage of children|minimum age of marriage|minimum age for marriage|marriage for girls and boys under the age |mutilation| fgm |honor crime|honor killing|honour crime|honour killing|breast ironing|harmful practices|harmful traditional practices|violence against women|domestic violence|family violence|violence in the family|gender-based violence|sexual violence|marital violence|partner violence|gender violence| rape|virginity|sexual abuse|femicide|polygam|sex selection|prenatal selection|prenatal genetic|dowry|bride kidnap|child bride|forced steril|istanbul convention| conversion therap|sexual harassment"

GBV_keywords_comb1a <- "forced|minimum|under the age of"
GBV_keywords_comb1b <- "marriage|pregnancy"

GBV_keywords_comb2a <- "girl|women|sexual|domestic|gender|marital|lgbt|same-sex"
GBV_keywords_comb2b <- "abuse|maltreatment|violence|harassment|intimidation"

GBV_keywords_comb3a <- "abort"
GBV_keywords_comb3b <- "select|forced|coercive"
```
#### Women's health
> This category is concerned with any health recommendation relating to women’s health, which was not linked to violence or SRHR, but rather addresses general health through a gendered lens. Recommendations classified here include the ratification of CEDAW.

May need to move this to end of the definitions, so that we can use the health_related definition and exclude those already included under SRHR or GBV
```{r}
women_keywords <- "cedaw|elimination of all forms of discrimination against women|menstrua"

women_keywords_comb1a <- "woman|women"
women_keywords_comb1b <- "mortality|health"
```
#### Child, and adolescent health
> This category covers issues relating to maternal, foetal, newborn, infant, child and adolescent health. Also included in this category are recommendations relating to the ratification of the Convention on the Rights of the Child (CRC) and its optional protocol regarding the sale of children.

```{r}
MCAH_keywords <- "newborn| infant|neonatal|new-born|breastfe|rights of the child| crc|sale of children|trafficking of children|child traffic|protection of women and children|trafficking of minor|exploitation of child|ill-treatment of girl|bacha bazi|violence against children"

MCAH_keywords_comb1a <- "child|infant|adolescent| teen|boy|girl"
MCAH_keywords_comb1b <- "death|mortality|pregnan|corporal punishment|health|traffic|violen|prostitut|smuggling|sexual abuse|exploitation"

# MCAH_keywords_comb2a <- "marriage"
# MCAH_keywords_comb2b <- "under the age of"
```

#### Essential medicines and health products
> The issue of essential medicines covers the availability of essential medicines within the context of functioning health systems. Ideally, essential medicines are available at all times in adequate amounts, in the appropriate dosage forms, with assured quality, and at a price the individual and the community can afford.

```{r}
essential_medicines_keywords <- "medicine|antibiotic|antimicrobial"
```


#### Disabilities and health
> Disability is an umbrella term for impairments, activity limitations and participation restrictions. This category includes any issue that relates to persons with disabilities and their right to access health services. Recommendations to ratify the Convention on the Rights of Persons with Disabilities fall in this category.

```{r}
disabilities_keywords <- "crpd|rights of persons with disabilities| deaf"

disabilities_keywords_comb1a <- "health"
disabilities_keywords_comb1b <- "disabil|disabled"
```

#### Health of LGBTI persons
> This category includes any issue related to lesbian, gay, bisexual, transgender, and intersex (LGBTI) persons and their right to access health services, including general health services but also specific LGBTI health needs, such as hormone therapy and sex reassignment surgery.

Difficult to tease this category out of SRHR based only on keywords - need to look in more details at the resulting recommendations and finetune it all.

Also need to refine for health-related rights vs rights in general
```{r}
LGBTI_keywords <- "lgbti|reassignment|intersex|same sex|same-sex"

LGBTI_keywords_comb1a <- "lgbt|lgtb| gay|lesbian|transgender|intersex|same sex|same-sex|gender"
LGBTI_keywords_comb1b <- "health|surgery|reassignment|forced"
```

#### HIV/AIDS and STIs
> This category is concerned with any recommendation relating to HIV/ AIDS and sexually transmitted infections (STIs), such as access to healthcare for persons with HIV/AIDS or STIs or prevention of the spread of HIV/AIDS and STIs.

```{r}
HIV_keywords <- "with hiv| hiv |hiv/aids|hiv-aids|hiv and aids|hiv-positive| hiv,| hiv-| sti | std | stds | stis | aids |sti,| std,| stds,| stis,| aids,|chlamydia|gonorrhea|syphilis|herpes|papillomavirus| hpv | hpv,|trichomoniasis|sexually transm|sexual transm|antiretroviral|anti-retroviral"

HIV_keywords_comb1a <- ""
HIV_keywords_comb1b <- ""
```
#### TB, Malaria, and NTDs
> This category includes any health issues related to tuberculosis, malaria or diseases such as rabies, leprosy, foodborne trematodiases, lymphatic filariasis, dracunculiasis (Guinea-worm disease), sleeping sickness, chikunguya, etc.

```{r}
TB_malaria_keywords <- "tuberculosis|malaria| tb | tb,"

NTD_keywords <- "buruli|chagas|dengue|chikungunya|dracunculiasis|echinococcosis|trematod|trypano|sleeping sickness|leishmaniasis|leprosy|leper|filariasis|mycetoma|chromoblastomycosis| noma |onchocerciasis|rabies|scabies|schistosomiasis|helminth|snakebite|snake|taenia|trachoma|yaws"
```
#### Immunization, vaccines and pharmaceutical products
> This category refers to any health issues related to immunization, vaccines and pharmaceutical products.

```{r}
vaccinations_keywords <- "vaccin|immunisation|immunization|medicine|antibiotic|antimicrobial|essential medicine"

vaccinations_keywords_comb1a <- ""
vaccinations_keywords_comb1b <- ""
```

#### WASH
> This category is concerned with access to safe or clean drinking water and basic standards of sanitation, including drinking-water quality management, water supply and sanitation monitoring, cholera surveillance and prevention, water and sanitation in different settings, and water resources management.

```{r}
WASH_keywords <- "sanitation|safe water|drinking water|drinkable water|affordable water|clean water|access to water|toilet|cholera"

WASH_keywords_comb1a <- ""
WASH_keywords_comb1b <- ""
```
#### Nutrition
> Recommendations relating to food were included when they referred to nutrition or diet-related concerns such as obesity, malnutrition or micronutrient supplementation.

```{r}
nutrition_keywords <- "hunger|food|nutrition|nutritious|diet|obesity|famine"

nutrition_keywords_comb1a <- ""
nutrition_keywords_comb1b <- ""
```

### Separate definitions

#### Occupational health
```{r}
occupational_health_keywords <- "health of workers|workers' health|occupational health"
```

#### Refugee and Migrant health
```{r}
migrant_health_keywords_comb1a <- "migrant|refugee|asylum"
migrant_health_keywords_comb1b <- "health"
```


#### Health of incarcerated persons
Recommendations related to the health and well-being of incarcerated persons. Only includes recommendations already linked to health
```{r}
incarcerated_keywords <- " prison|imprison|detention|detaine|incarcer"
```


#### Maternal health
```{r}
maternal_keywords <- "maternal|pregnan|midwif|midwiv|antenatal|prenatal|postnatal|birth atten|childbirth|child birth|child-birth|stillbirth|eclampsia|obstetric|miscarr|intrapartum|postpartum|post-partum|breastfe|safe motherhood|protecting maternity|health of the mother|life of the mother|mothers' health|mother and child health|mothers and newborns to full health care|care for children and mothers"

maternal_keywords_comb1a <- "motherhood"
maternal_keywords_comb1b <- "health"

NOT_maternal_keywords <- "miscarriage of justice|miscarriages of justice"

```

#### Abortion

#### Family planning/Contraception
```{r}
contraception_keywords <- "contracept|birth control| iud |condom|family planning"

contraception_keywords_comb1a <- "unintended pregnan"
contraception_keywords_comb1b <- "prevent|protect"
```


### Some terms to watch out for and consider for refining false positives:

traffic / trafficking


### Apply the definitions

### Theme labels
```{r}
source(here("code", "theme_labels.R"))
```

### UHRI datasat
```{r}
sdg_data <- uhri_upr |> 
  # Health systems and services
  mutate(
    health_systems = factor(case_when(
      str_detect(tolower(text_2), hss_keywords) ~ "Health systems and services",
      # Combo 1
      str_detect(tolower(text_2), hss_keywords_comb1a)&
        str_detect(tolower(text_2), hss_keywords_comb1b)~ "Health systems and services",
      
      # Combo 1
      str_detect(tolower(text_2), hss_keywords_comb2a)&
        str_detect(tolower(text_2), hss_keywords_comb2b)&
        str_detect(tolower(text_2), hss_keywords_comb2c) ~ "Health systems and services",
      
      .default = "Other"
    ), levels = c("Other", "Health systems and services"))
  ) |> 
  
  # Health security, emergencies, and disaster relief
  mutate(
    emergencies = factor(case_when(
      str_detect(tolower(text_2), emergencies_keywords) ~ "Health security, emergencies, and disaster relief",
      # Combo 1
      str_detect(tolower(text_2), emergencies_keywords_comb1a)&
        str_detect(tolower(text_2), emergencies_keywords_comb1b)~ "Health security, emergencies, and disaster relief",
      
      .default = "Other"
    ), levels = c("Other", "Health security, emergencies, and disaster relief"))
  ) |> 
  
  # NCDs
  mutate(
    ncd = factor(case_when(
      str_detect(tolower(text_2), ncd_keywords) ~ "NCDs",
      # Combo 1
      str_detect(tolower(text_2), ncd_keywords_comb1a)&
        str_detect(tolower(text_2), ncd_keywords_comb1b)~ "NCDs",
      
      # Combo 2
      str_detect(tolower(text_2), ncd_keywords_comb2a)&
        str_detect(tolower(text_2), ncd_keywords_comb2b)~ "NCDs",
      
      .default = "Other"
    ), levels = c("Other", "NCDs"))
  ) |> 
  
  # Communicable diseases
  mutate(
    communicable = factor(case_when(
      str_detect(tolower(text_2), communicable_keywords) ~ "Communicable diseases",
      # Combo 1
      str_detect(tolower(text_2), communicable_keywords_comb1a)&
        # Indicate "!" to exclude these terms
        !str_detect(tolower(text_2), communicable_keywords_comb1b) ~ "Communicable diseases",
      
      .default = "Other"
    ), levels = c("Other", "Communicable diseases"))
  ) |> 
  
  # SRHR
  mutate(
    SRHR = factor(case_when(
      str_detect(tolower(text_2), srhr_keywords) ~ "SRHR",
      # Combo 1
      str_detect(tolower(text_2), srhr_keywords_comb1a)&
        str_detect(tolower(text_2), srhr_keywords_comb1b)~ "SRHR",
      
      # Combo 2
      str_detect(tolower(text_2), srhr_keywords_comb2a)&
        str_detect(tolower(text_2), srhr_keywords_comb2b)~ "SRHR",
      
      .default = "Other"
    ), levels = c("Other", "SRHR"))
  ) |> 
  
  # Mental health
  mutate(
    mental_health = factor(case_when(
      str_detect(tolower(text_2), mental_keywords) ~ "Mental health",
      # Combo 1
      str_detect(tolower(text_2), mental_keywords_comb1a)&
        str_detect(tolower(text_2), mental_keywords_comb1b)~ "Mental health",
      
      .default = "Other"
    ), levels = c("Other", "Mental health"))
  ) |> 
  
  # Social and economic determinants
  
  # GBV
  mutate(
    GBV = factor(case_when(
      str_detect(tolower(text_2), GBV_keywords) ~ "GBV",
      # Combo 1
      str_detect(tolower(text_2), GBV_keywords_comb1a)&
        str_detect(tolower(text_2), GBV_keywords_comb1b)~ "GBV",
      # Combo 2
      str_detect(tolower(text_2), GBV_keywords_comb2a)&
        str_detect(tolower(text_2), GBV_keywords_comb2b)~ "GBV",
      
      # Combo 3
      str_detect(tolower(text_2), GBV_keywords_comb3a)&
        str_detect(tolower(text_2), GBV_keywords_comb3b)~ "GBV",
      
      .default = "Other"
    ), levels = c("Other", "GBV"))
  ) |> 
  
  # Maternal, child, and adolescent health (MCAH)
  mutate(
    MCAH = factor(case_when(
      str_detect(tolower(text_2), MCAH_keywords) ~ "MCAH",
      # Combo 1
      str_detect(tolower(text_2), MCAH_keywords_comb1a)&
        str_detect(tolower(text_2), MCAH_keywords_comb1b)~ "MCAH",
      
      # # Combo 2
      # str_detect(tolower(text_2), MCAH_keywords_comb2a)&
      #   str_detect(tolower(text_2), MCAH_keywords_comb2b)~ "MCAH",
      
      .default = "Other"
    ), levels = c("Other", "MCAH"))
  ) |> 
  
  # # Essential medicines and health products
  # mutate(
  #   essential_medicines = factor(case_when(
  #     str_detect(tolower(text_2), essential_medicines_keywords) ~ "Essential medicines",
  #     # # Combo 1
  #     # str_detect(tolower(text_2), disabilities_keywords_comb1a)&
  #     #   str_detect(tolower(text_2), disabilities_keywords_comb1b)~ "Essential medicines",
  #     .default = "Other"), 
  #     levels = c("Other", "essential_medicines_keywords"))
  # ) |> 
  
  # Disabilities and health
  mutate(
    disabilities = factor(case_when(
      str_detect(tolower(text_2), disabilities_keywords) ~ "Disabilities and health",
      # Combo 1
      str_detect(tolower(text_2), disabilities_keywords_comb1a)&
        str_detect(tolower(text_2), disabilities_keywords_comb1b)~ "Disabilities and health",
      .default = "Other"), 
      levels = c("Other", "Disabilities and health"))
  ) |> 
  
  # Health of LGBTI persons
  mutate(
    LGBTI = factor(case_when(
      # str_detect(tolower(text_2), LGBTI_keywords) ~ "LGBTI",
      # Combo 1
      str_detect(tolower(text_2), LGBTI_keywords_comb1a)&
        str_detect(tolower(text_2), LGBTI_keywords_comb1b)~ "LGBTI",
      .default = "Other"), 
      levels = c("Other", "LGBTI"))
  ) |> 
  
  # HIV/AIDS and STIs
  mutate(
    HIV = factor(case_when(
      str_detect(tolower(text_2), HIV_keywords) ~ "HIV",
      .default = "Other"), 
      levels = c("Other", "HIV"))
  ) |> 
  
  # TB, malaria, and NTDs
  mutate(
    TB_malaria = factor(case_when(
      str_detect(tolower(text_2), TB_malaria_keywords) ~ "TB and Malaria",
      .default = "Other"), 
      levels = c("Other", "TB and Malaria")),
    NTD = factor(case_when(
      str_detect(tolower(text_2), NTD_keywords) ~ "NTD",
      .default = "Other"), 
      levels = c("Other", "NTD")),
    TB_malaria_NTD = factor(case_when(
      if_any(c(TB_malaria, NTD), ~ .x != "Other") ~ "TB, malaria, and NTDs",
      .default = "Other"),
      levels = c("Other", "TB, malaria, and NTDs"))
  ) |> 
  
  # Vaccinations
  mutate(
    vaccinations = factor(case_when(
      str_detect(tolower(text_2), vaccinations_keywords) ~ "Vaccinations",
      .default = "Other"), 
      levels = c("Other", "Vaccinations"))
  ) |> 
  
  # WASH
  mutate(
    WASH = factor(case_when(
      str_detect(tolower(text_2), WASH_keywords) ~ "WASH",
      .default = "Other"), 
      levels = c("Other", "WASH"))
  ) |> 
  
  # Nutrition
  mutate(
    nutrition = factor(case_when(
      str_detect(tolower(text_2), nutrition_keywords) ~ "Nutrition",
      # Combo 1
      # str_detect(tolower(text_2), nutrition_keywords_comb1a)&
      #   str_detect(tolower(text_2), nutrition_keywords_comb1b)~ "Nutrition",
      
      .default = "Other"
    ), levels = c("Other", "Nutrition"))
  ) |> 
  
  # Right to health
  mutate(
    health_related = factor(case_when(
      
      # recommendations that fall under other health-related themes
      if_any(c(health_systems:nutrition), ~ .x != "Other") ~ "Health-related",
      
      # Based on keywords
      str_detect(tolower(text_2), NOT_health_keywords) ~ "Other",
      str_detect(tolower(text_2), health_keywords) ~ "Health-related",
      # str_detect(tolower(themes), "health|sanitation") ~ "Health-related",
      # str_detect(targets, "5.6|16.1|16.2")
      # # Combo 1
      # str_detect(tolower(text_2), keywords_comb1a)&
      #   str_detect(tolower(text_2), keywords_comb1b)~ "Health-related",
      # # Combo 2
      # str_detect(tolower(text_2), keywords_comb2a)&
      #   str_detect(tolower(text_2), keywords_comb2b)~ "Health-related",
      # # Combo 3
      # str_detect(tolower(text_2), keywords_comb3a)&
      #   str_detect(tolower(text_2), keywords_comb3b)~ "Health-related",
      
      .default = "Other"
    ), levels = c("Other", "Health-related"))
  ) |> 
  
  # Women's health
  mutate(
    women = NA # (re-defined later on)
  ) |> 
  
  # Abortion
  mutate(
    abortion = factor(case_when(
      str_detect(tolower(text_2), "abortion|foetal impairment|fetal impairment") ~ "Abortion",
      # Combo 1
      str_detect(tolower(text_2), "interrup|terminat")&
        str_detect(tolower(text_2), "pregnan")~ "Abortion",
      .default = "Other"),
      levels = c("Other", "Abortion"))) |> 
  
  # Family planning/Contraception
  mutate(
    contraception = factor(case_when(
      str_detect(tolower(text_2), contraception_keywords) ~ "Contraception",
      # Combo 1
      str_detect(tolower(text_2), contraception_keywords_comb1a)&
        str_detect(tolower(text_2), contraception_keywords_comb1b)~ "Contraception",
      .default = "Other"), 
      levels = c("Other", "Contraception"))
  ) |> 
  
  # Incarcerated persons
  mutate(
    incarcerated = factor(case_when(
      health_related == "Health-related" & 
        str_detect(tolower(text_2), incarcerated_keywords) ~ "Incarcerated",
      .default = "Other"),
      levels = c("Other", "Incarcerated"))) |> 
  
  # Occupational health
  mutate(
    occupational_health = factor(case_when(
      str_detect(tolower(text_2), occupational_health_keywords) ~ "Occupational health",
      # # Combo 1
      # str_detect(tolower(text_2), occupational_health_keywords_comb1a)&
      #   str_detect(tolower(text_2), occupational_health_keywords_comb1b)~ "Occupational health",
      .default = "Other"), 
      levels = c("Other", "Occupational health"))
  ) |> 
  
  # Migrant health
  mutate(
    migrant_health = factor(case_when(
      # str_detect(tolower(text_2), migrant_health_keywords) ~ "Migrant health",
      # Combo 1
      str_detect(tolower(text_2), migrant_health_keywords_comb1a)&
        str_detect(tolower(text_2), migrant_health_keywords_comb1b)~ "Migrant health",
      .default = "Other"), 
      levels = c("Other", "Migrant health"))
  ) |> 
  
  # Maternal health
  mutate(
    maternal_health = factor(case_when(
      str_detect(tolower(text_2), NOT_maternal_keywords) ~ "Other",
      str_detect(tolower(text_2), maternal_keywords) ~ "Maternal health",
      # Combo 1
      str_detect(tolower(text_2), maternal_keywords_comb1a)&
        str_detect(tolower(text_2), maternal_keywords_comb1b) ~ "Maternal health",
      .default = "Other"
    ), levels = c("Other", "Maternal health")
    )
  ) |> 
  
  # Update overall Right to Health classification
  mutate(
    health_related = factor(case_when(
      if_any(c(abortion, maternal_health, contraception,
               occupational_health, migrant_health), ~ .x != "Other") ~ "Health-related",
      .default = health_related),
      levels = c("Other", "Health-related"))
  )|> 
  
  # Update Health systems and services
  mutate(
    health_systems = factor(case_when(
      health_related == "Health-related" & 
        str_detect(tolower(text_2), hss_other_keywords) ~ "Health systems and services",
      .default = health_systems),
      levels = c("Other", "Health systems and services"))
  ) |> 
  
  # Update women's health
  mutate(
    women = factor(case_when(
      str_detect(tolower(text_2), women_keywords) ~ "Women's health",
      # Exclude recommendations related to GBV or SRHR
      if_any(c(GBV, SRHR, maternal_health), ~ .x != "Other") ~ "Other",
      # Combo 1
      str_detect(tolower(text_2), women_keywords_comb1a)&
        str_detect(tolower(text_2), women_keywords_comb1b)~ "Women's health",
      # Assign health-related topics that relate to women
      health_related == "Health-related" & 
        str_detect(tolower(text_2), "women|woman") ~ "Women's health",
      .default = "Other"
    ), levels = c("Other", "Women's health"))
  ) |> 
  
  # # Update SRHR
  # mutate(
  #   SRHR = factor(case_when(
  #     if_any(c(abortion, contraception), ~ .x != "Other") ~ "SRHR",
  #     .default = SRHR),
  #     levels = c("Other", "SRHR"))
  # )|> 
  
  # Update overall Right to Health classification
  mutate(
    health_related = factor(case_when(
      if_any(c(women), ~ .x != "Other") ~ "Health-related",
      .default = health_related),
      levels = c("Other", "Health-related"))
  )|> 
  
  mutate(
    other_health_related = factor(case_when(
      if_all(c(health_systems:maternal_health, -health_related), ~ .x == "Other") & health_related != "Other" ~ "Other health-related",
      .default = "Other"),
      levels = c("Other", "Other health-related"))
  )|> 
  
  # Reorder columns
  relocate(health_systems:maternal_health, other_health_related, .after = text) |> 
  relocate(health_related, .after = text)

saveRDS(sdg_data, file = here("output", "UHRI_UPR_enhanced.rds"))
```

```{r}
sdg_data <- sdg_data |>
  mutate(response_upr = fct_recode(response_upr, 
                                   "Noted" = "Partially supported"),
         response_upr = fct_relevel(response_upr, "Noted")) |> 
  droplevels()
```


### Number of recommendations
```{r}
#| message: false
#| warning: false
#| fig-width: 5
#| fig-height: 4
sdg_data |>
  # filter(type == "Recommendations") |> 
  group_by(cycle) |>
  summarise(
    n = n(),
    n_SuR = n_distinct(state_under_review),
    # n_rec_state = n_distinct(recommending_state_upr),
    n_health = sum(health_related=="Health-related"),
    n_not_health = sum(health_related == "Other")
  ) |>
  ggplot(aes(x = cycle, y = n))+
  geom_col()+
  labs(y = "Number of recommendations", x = "UPR Cycle", 
       title = "Number of recommendations received by States",
       fill = NULL)+
  geom_text(aes(label = format(n, big.mark = ","), y = n, vjust =-0.2), size = 4)+
  theme_bw()+
  theme(panel.grid = element_blank(),
        legend.position = "inside",
        legend.position.inside = c(0,1),
        legend.justification = c("left", "top"),
        legend.background = element_blank())
```


Median number of recommendations per cycle
```{r}
#| fig-height: 5.5
#| fig-width: 7
#| message: false
sdg_data |> 
  group_by(cycle, state_under_review) |> 
  count(health_related, .drop = FALSE) |> group_by(cycle, health_related) |> mutate(med_n = median(n)) |> 
  select(cycle, health_related, med_n) |> distinct() |> 
  group_by(cycle) |> 
  mutate(n_tot = sum(med_n),
         perc = (med_n/n_tot)*100,
         perc = case_when(health_related == "Other" ~ "",
                          .default = paste0(sprintf("%1.1f",perc), "%"))) |> 
  ggplot(aes(x = cycle, y = med_n, fill = health_related))+
  geom_bar(stat = "identity")+
  labs(y = "Median number of recommendations", x = "UPR Cycle", 
       title = "Median number of recommendations received by States, per UPR cycle",
       fill = NULL)+
  geom_text(aes(label = perc), position = position_stack(vjust = 0.5), size = 5)+
  geom_text(aes(label = sprintf("%1.0f", n_tot), y = n_tot, vjust =-0.2), size = 5)+
  theme_bw()+
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(size = 18),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 16),
        legend.position = "inside",
        legend.position.inside = c(0,1),
        legend.justification = c("left", "top"),
        legend.text = element_text(size = 14),
        legend.background = element_blank())
```

Boxplots of number of recommendations
```{r}
#| message: false
#| warning: false
sdg_data |> 
  group_by(cycle, state_under_review) |> 
  count(health_related, .drop = FALSE) |> 
  ungroup() |> 
  ggplot(aes(x = cycle, y = n, fill =  health_related))+ 
  labs(title = "Boxplots of the number of recommendations received by each state",
       y = "Number of recommendations", color = NULL, fill = NULL, x = NULL)+
  geom_jitter(aes(color = health_related), position = position_jitterdodge())+
  geom_boxplot(outlier.shape = NA, alpha = 0.8, staplewidth = 0.5)+
  # stat_summary(
  #   fun = median, 
  #   geom = "text", 
  #   aes(label = after_stat(y)), 
  #   position = position_dodge(0.75),
  #   hjust = -2,
  #   # nudge_x = -0.1,
  #   # color = "black",
  #   size = 3.5
  # ) +
  theme_bw()+
  theme(legend.position = "bottom", panel.grid.major.x = element_blank())
```

#### By theme
```{r, message=FALSE}
a_1 <- sdg_data |>
  select(cycle, health_related:other_health_related, response_upr) |>
  group_by(response_upr) |>
  summarise(across(c(health_related:other_health_related), ~ sum(.x != "Other"))) |>
  ungroup() |>
  filter(response_upr %in% c("Supported", "Noted")) |>
  pivot_longer(
    cols = health_related:other_health_related,
    names_to = "theme",
    values_to = "n"
  )

a_2 <- sdg_data |>
  select(cycle, health_related:other_health_related, response_upr) |>
  group_by(response_upr) |>
  summarise(across(c(health_related:other_health_related), ~ sum(.x == "Other"))) |>
  ungroup() |>
  filter(response_upr %in% c("Supported", "Noted")) |>
  pivot_longer(
    cols = health_related:other_health_related,
    names_to = "theme",
    values_to = "n_other"
  )

a <- left_join(a_1, a_2) |>
  group_by(theme) |>
  mutate(
    n_tot = sum(n) + sum(n_other),
    n_tot_theme = sum(n)
  ) |>
  mutate(
    perc = n / n_tot * 100,
    perc_theme = n_tot_theme / n_tot * 100
  ) |>
  group_by(theme) |>
  mutate(
    n_sup = paste0("(", sprintf("%1.0f", n / sum(n) * 100), "%)"),
    n_sup = case_when(n_tot_theme == 0 ~ "", .default = n_sup)
  ) |>
  ungroup() |>
  filter(!theme %in% c("health_related", "TB_malaria", "NTD")) |>
  left_join(theme_labels, by = c("theme" = "variable")) |>
  arrange(-n_tot_theme) |>
  mutate(
    theme_label = case_when(is.na(theme_label) ~ theme, .default = theme_label),
    # theme_label = str_wrap(theme_label, width = 40),
    theme_label = fct_inorder(theme_label)
  )


max_a <- max(a$perc_theme, na.rm = TRUE)
a |>
  ggplot(aes(x = perc, y = fct_rev(theme_label))) +
  geom_col(aes(fill = response_upr)) +
  labs(
    x = "Proportion of all recommendations (%)", y = NULL,
    fill = "State's response",
    title = paste0("All health-related recommendations of the UPR"),
    caption = "*Numbers after the bars indicate N (% supported)"
  ) +
  theme_classic() +
  scale_x_continuous(
    labels = function(x) paste0(x, "%"),
    # limits = c(0, max_a + 2),
    expand = expansion(mult = c(0, 0.15))
  ) +
  coord_cartesian(clip = "off")+
  theme(
    plot.margin = margin(l=2,t=2,b=2, r = 30, unit = "pt"),
    legend.position = c(0.99, 0.01),
    legend.justification = c("right", "bottom"),
    legend.frame = element_rect(color = "black"),
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 11),
    legend.background = element_rect(fill = "transparent"),
    axis.text.y = element_text(size = 9),
    axis.text.x = element_text(size = 10),
    plot.title = element_text(hjust = 0.5,face = "bold"),
    plot.caption = element_text(size = 11),
    axis.title.y = element_blank(),
    plot.title.position = "plot",
    panel.grid = element_blank()
  ) +
  geom_text(
    data = a |> filter(response_upr == "Supported"),
    aes(label = paste0(n_tot_theme, " ", n_sup), x = perc_theme),
    hjust = -0.05,
    size = 3, color = "#1c164d"
    # vjust = 0.25
  )
```

```{r}
#| fig-height: 9.5
#| fig-width: 8
#| message: false

a_1<-sdg_data |> 
  # filter(state_under_review == "Nigeria") |> 
  select(cycle, state_under_review, health_related:other_health_related, response_upr) |> 
  # mutate(across(c(health_related:other_health_related), ~ .x != "Other")) |> 
  group_by(cycle, response_upr) |> 
  summarise(across(c(health_related:other_health_related), ~ sum(.x !="Other"))) |> 
  ungroup() |> 
  filter(response_upr %in% c("Supported", "Noted")) |>
  pivot_longer(cols = health_related:other_health_related, 
               names_to = "theme", 
               values_to = "n"
  )

a_2<-sdg_data |> 
  # filter(state_under_review == "Nigeria") |> 
  select(cycle, state_under_review, health_related:other_health_related, response_upr) |> 
  # mutate(across(c(health_related:other_health_related), ~ .x != "Other")) |> 
  group_by(cycle, response_upr) |> 
  summarise(across(c(health_related:other_health_related), ~ sum(.x =="Other"))) |> 
  ungroup() |> 
  filter(response_upr %in% c("Supported", "Noted")) |>
  pivot_longer(cols = health_related:other_health_related, 
               names_to = "theme", 
               values_to = "n_other"
  )

a_3<-sdg_data |> 
  # filter(state_under_review == "Nigeria") |> 
  group_by(cycle) |> 
  summarise(health_n = sum(health_related !="Other")) |> 
  ungroup()

a <- left_join(a_1,a_2) |> 
  left_join(a_3) |> 
  mutate(cycle = fct_recode(cycle, "1"="Cycle 1", "2"="Cycle 2", "3"="Cycle 3", "4"="Cycle 4")) |> 
  group_by(cycle, theme) |> 
  mutate(n_tot = sum(n)+sum(n_other)) |> 
  mutate(n_tot_theme = sum(n)) |> 
  mutate(perc = n/n_tot*100,
         perc_theme = n_tot_theme/n_tot*100,
         theme_perc_health = n_tot_theme/health_n*100) |> 
  group_by(cycle, theme) |> 
  mutate(n_sup = paste0("(", sprintf("%1.0f", n/sum(n)*100), "%)"),
         n_sup = case_when(n_tot_theme == 0 ~ "(NA)", .default = n_sup)) |> 
  # mutate(n_sup = case_when(
  #   response_upr == "Noted/Other" ~ "",
  #   response_upr == "Supported" ~ paste0("(", sprintf("%1.0f", n/sum(n)*100), "%)"),
  #   .default = NA
  # )) |> 
  ungroup() |> 
  filter(!theme %in% c(
    "health_related"
    # , "abortion"
    , "TB_malaria", "NTD"
    # , "TB_malaria_NTD"
  )) |> 
  left_join(theme_labels, join_by(theme == variable)) |> 
  arrange(cycle, -n_tot_theme) |> 
  mutate(theme_label = case_when(is.na(theme_label) ~ theme, .default = theme_label),
         theme_label = fct_inorder(theme_label))

max_a <- max(a$perc_theme)
theme_plot <- a |> 
  # mutate(n_tot_theme = case_when(response_upr!="Supported" ~ "", 
  #                                .default = as.character(n_tot_theme))) |> 
  ggplot(aes(x = perc, y = fct_rev(cycle)))+
  geom_col(aes(fill = response_upr))+
  facet_grid(
    rows = vars(theme_label), switch = "y",
    labeller = labeller(theme_label = label_wrap_gen(40))
  )+
  labs(x = "Proportion of all recommendations within the respective UPR cycle (%)", y = NULL,
       fill = "State's response",
       title = "Health-related recommendations in the first four cycles\nof the Universal Periodic Review",
       caption = "*Numbers after the bars indicate N (% supported)")+
  # theme_bw()+
  theme_classic()+
  # scale_y_discrete(expand = c(0.1, 0))+
  scale_x_continuous(labels = function(x) paste0(x, "%"), 
                     limits = c(0,max_a+2), 
                     # sec.axis = dup_axis(name = NULL),
                     expand = expansion(mult = c(0, 0.05)) # 0 exactly on axis
  )+
  theme(
    legend.position = "inside",
    legend.position.inside = c(0.75,0.5),
    legend.justification = c("right", "center"),
    legend.frame = element_rect(color = "black"),
    axis.text.y = element_text(size = 6), 
    plot.title = element_text(hjust = 0.5),
    strip.placement = "outside",
    strip.text.y.left = element_text(angle = 0, vjust = 1, size = 8),
    # strip.background = element_blank()
    panel.grid = element_blank()
  )+
  geom_text(
    data = a |> filter(response_upr == "Supported"),
    aes(label = paste0(n_tot_theme, " ", n_sup), x = perc_theme),
    hjust = -0.15, size = 2.5 , vjust=0.25
  ); theme_plot

# ggsave(theme_plot, filename = "theme_plot.png", height = 9.5, width = 8)
# 
# a |> 
#   filter(theme %in% c("SRHR", "abortion"), response_upr == "Supported") |> 
#   select(cycle, theme, n_tot_theme, n_tot, perc_theme, health_n, theme_perc_health, n_sup) |> 
#   mutate(n_sup = str_remove_all(n_sup, "\\(|\\%|\\)"),
#          n_sup = as.numeric(n_sup)/100) |> 
#   arrange(theme, cycle) |> 
#   mutate(perc_theme = perc_theme/100, theme_perc_health = theme_perc_health/100) |> 
#   rename(
#     "UPR cycle" = cycle,
# "Theme" = theme,
# "Recommendations in theme" = n_tot_theme,
# "Total recommendations" = n_tot,
# "Health recommendations" = health_n,
# "% of total recommendations" = perc_theme,
# "% of health recommendations" = theme_perc_health,
# "Percent of recommendations supported" = n_sup
# )

```

```{r}
#| fig-width: 13
#| fig-height: 5
p1<-a |> 
  filter(!theme %in% c("health_related")) |> 
  filter(theme %in% c(
    # "GBV"
    "maternal_health",
    "abortion",
    "contraception",
    "SRHR"
  )) |> 
  # group_by(cycle, theme) |> 
  ggplot(aes(x = cycle, y = perc, fill = response_upr))+
  # scale_fill_manual(values = c("Noted/Other" = "#FFC300", "Supported" = "#5669f1"))+
  labs(x = "UPR cycle", y = "% of total recommendations", 
       fill = "State's response",
       title = "Percent of total UPR recommendations by theme\nSupported % in brackets")+
  geom_col()+
  geom_text(aes(label = n_sup), size = 5, hjust = 0.5, color = "grey20", 
            position = position_stack(vjust = 0.5))+
  # geom_text(aes(label = n_sup, y = if_else(perc<1,1.6,perc/2)), size = 5, color = "white")+
  # geom_text(aes(label = sprintf("%1.0f", n_tot_theme), y = sum(perc), vjust =-0.2), size = 5)+
  facet_wrap(.~theme,
             nrow = 1,
             labeller = as_labeller(c("abortion" = "Abortion",
                                      "GBV" = "Gender-based violence",
                                      "contraception" = "Contraception/Family planning",
                                      "maternal_health" = "Maternal health",
                                      "SRHR" = "SRHR")))+
  theme_bw()+
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.text.x = element_text(size = 12),
        axis.title.x = element_text(size=16),
        axis.title.y = element_text(size = 16),
        strip.text = element_text(size = 16),
        plot.title = element_text(size = 16),
        legend.position = "inside",
        legend.position.inside = c(0,1),
        legend.justification = c("left", "top"),
        legend.title = element_text(size = 15),
        legend.text = element_text(size = 14),
        legend.background = element_blank())+
  scale_y_continuous(sec.axis = dup_axis(name = NULL)
                     # , limits = c(0,13)
  );p1
# ggsave("recommendations_by_theme.png", p1, width = 8, height = 4.5)
# ggsave("recommendations_by_theme_gbv.png", p1, width = 9, height = 4)

```

#### Recommending countries
See <https://mattherman.info/blog/fix-facet-width/>
```{r, message=FALSE}
c<-sdg_data |> 
  # filter(state_under_review == "Brazil") |> 
  # filter(maternal_health != "Other") |> 
  filter(response_upr == "Supported") |> 
  filter(if_any(any_of(c(theme_labels$variable)), ~ .x != "Other")) |> 
  # select(cycle, recommending_state_upr) |> 
  separate_longer_delim(cols = c(recommending_state_upr), delim="-") |> 
  mutate(recommending_state_upr = str_trim(recommending_state_upr)) |> 
  group_by(state_under_review, cycle, recommending_state_upr) |> count(across(any_of(theme_labels$variable))) |> 
  pivot_longer(cols = !c(state_under_review,cycle, recommending_state_upr, n),
               names_to = "variable",
               values_to = "theme") |> 
  filter(theme!="Other") |> 
  left_join(theme_labels) |> 
  group_by(state_under_review, cycle, recommending_state_upr, variable) |> 
  mutate(n=sum(n)) |> 
  ungroup() |> 
  distinct()

c_plot <- c |> 
  # filter(state_under_review == "Tonga") |>
  filter(variable %in% c("abortion", "maternal_health", "contraception")) |> 
  select(-cycle, -state_under_review) |> 
  group_by(recommending_state_upr, theme) |> 
  mutate(n=sum(n)) |> 
  ungroup() |> 
  distinct() |> 
  group_by(recommending_state_upr) |> 
  mutate(n_tot = sum(n)) |> 
  ungroup() |> 
  arrange(-n_tot) |>
  # arrange(theme, -n) |> 
  # group_by(theme) |> 
  # slice_head(n = 10) |>
  ungroup()

ccp <- c_plot |> 
  arrange(theme, -n) |> 
  group_by(theme) |> 
  # select(recommending_state_upr, n_tot) |> distinct() |>
  # arrange(-n_tot) |> 
  slice_head(n=15)

c_plot |> 
  filter(recommending_state_upr %in% c(ccp |> pull(recommending_state_upr))) |> 
  # ggplot(aes(x= tidytext::reorder_within(recommending_state_upr, n, theme_label), y=n, fill=theme_label))+
  ggplot(aes(x= reorder(recommending_state_upr, n_tot), y=n,fill=theme_label))+
  geom_col(alpha = 0.8, width = 0.85)+
  # scale_fill_brewer(palette = "Dark2") +
  scale_fill_manual(values = c(
    "Maternal health" = "#7570b3",
    "Family planning" = "#1b9e77",
    "Abortion" = "#d95f02"
  ))+
  scale_y_continuous(
    expand = c(0, 0.1), 
    # breaks = c(0:max(c_plot$n_tot))
  ) +
  tidytext::scale_x_reordered() +
  coord_flip()+
  theme_minimal() +
  # facet_wrap(.~theme_label, scales = "free_y", strip.position="left", space = "free_y",
  #            labeller = labeller(theme_label = label_wrap_gen(10)))+
  guides(fill=guide_legend(reverse=TRUE, nrow=1))+
  theme(
    strip.text.y = element_text(angle = 270, face = "bold"),
    panel.grid.minor = element_blank(),
    strip.placement = "outside",
    panel.grid.major.y = element_blank(),
    axis.text = element_text(size=11),
    axis.title = element_text(size=11),
    legend.text = element_text(size=12),
    # legend.position = c(0.99, 0.01),
    legend.position = "top",
    legend.justification = c("left", "bottom"),
    legend.margin = margin(0,0,0,0), 
    legend.box.margin = margin(0,0,0,0), 
    legend.box.spacing = unit(0,"pt"),
    aspect.ratio = 0.04*n_distinct(ccp$recommending_state_upr),
    legend.background = element_blank())+
  labs(y="Number of issued recommendations that were supported", x=NULL,
       fill=NULL)
```

```{r, message=FALSE}
d <- c |> 
  filter(variable %in% c("abortion", "maternal_health", "contraception")) |> 
  group_by(recommending_state_upr, variable) |> 
  summarise(n = sum(n)) |>
  arrange(variable, -n) |> 
  group_by(variable) |> slice_head(n=15) |> 
  ungroup() |> 
  left_join(theme_labels)

d |> 
  mutate(recommending_state_upr = str_wrap(recommending_state_upr,20)) |> 
  ggplot(aes(x=tidytext::reorder_within(recommending_state_upr, n, theme_label), y=n))+
  geom_col()+
  tidytext::scale_x_reordered() +
  coord_flip()+
  theme_classic()+
  facet_wrap(.~theme_label, scales = "free_y")+
  # guides(fill="none")+
  labs(y="Number of issued recommendations that were supported", 
       x=NULL)




#   arrange(cycle, variable, -n) |>
#   group_by(cycle, variable) |> 
#   # slice_head(n = 10) #|> 
# ggplot(aes(x= tidytext::reorder_within(recommending_state_upr, n_tot, theme_label), y=n_tot))+
#   geom_col()+
#   tidytext::scale_x_reordered() +
#   coord_flip()+
#   theme_classic()+
#   facet_wrap(.~theme_label, scales = "free_y")+
#   # guides(fill="none")+
#   labs(y="Number of issued recommendations that were supported", x=NULL)
```